### 2015-05-25 rapport/component.tex
rapport_25_05

%-------------------------------------
\subsection{A primitive component model}

%-------------------------------------
\subsection{Control components}

%-------------------------------------
\subsection{Transformation to a parallel assembly}