### 2015-06-02 rapport/msp.tex
rapport_02_06_15

In this section we argue that the graphs $\Gamma_{hybrid}$ or $\Gamma_{task}$, previously defined, on which an approximation will be defined, are minimal series-parallel graphs. For this reason, the structure of those graphs can be represented as binary trees of parallel and series compositions of sub-graphs, also called \emph{binary decomposition tree}~\cite{Valdes:1979:RSP:800135.804393}. First, the needed definitions on series-parallel graphs will be given...

%--------------------
\subsection{GSP and MSP classes}
A vertex $v$ of a DAG $G$ is a \emph{source} if no edge of $G$ enters $v$. Similarly, a vertex $v$ is a \emph{sink} if no edge of $G$ leaves $v$. In 1982, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have defined the class of minimal series-parallel DAGs (MSP).

\begin{mydef}Minimal Series Parallel
\begin{itemize}
\item The DAG having a single vertex and no edges is MSP.
\item If $G_1=(V_1,E_1)$ and $G_2=(V_2,E_2)$ are two MSP DAGs, so is either of the DAGs constructed by the following operations:
\begin{itemize}
\item Parallel composition: $G_p=(V_1\cup V_2,E_1\cup E_2)$.
\item Series composition: $G_s=(V_1\cup V_2,E_1\cup E_2\cup (N_1 \times R_2))$, where $N_1$ is the set of sinks of $G_1$ and $R_2$ is the set of sources of $G_2$.
\end{itemize}
\end{itemize}
\end{mydef}

\begin{mydef}
A DAG is \emph{General Series Parallel} (GSP) if and only if its transitive reduction is a MSP DAG.
\end{mydef}

A \emph{binary decomposition tree} is a tree having a leaf for each vertex of the MSP DAG it represents, and whose internal nodes are labelled $S$ or $P$ to indicate respectively the series or parallel composition of the MSP sub-DAGs represented by the subtrees rooted at $S$ or $P$. Figures~\ref{fig:gsp}, ~\ref{fig:msp} and~\ref{fig:t} respectively give an example of a GSP DAG, its transitive reduction which is MSP, and its tree decomposition.

\begin{figure}[h!]
\begin{center}
\subfloat[][\label{fig:gsp}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (a) at (0,0) {$a$};
   \node[] (b) at (1,0) {$b$};
   \node[] (c) at (2,0.5) {$c$};
   \node[] (d) at (3,0.5) {$d$};
   \node[] (e) at (2,-0.5) {$e$};
   \node[] (f) at (3,-0.5) {$f$};
   \node[] (g) at (4,0) {$g$};
   \node[] (h) at (1,-1) {$h$};
   \node[] (i) at (2,-1) {$i$};
   \node[] (j) at (3,-1) {$j$};

  \path[->]
    (a) edge node {} (b)
        edge [bend left=50] node [swap] {} (d)
        edge node {} (h)
    (b) edge node {} (c)
        edge node {} (e)
    (c) edge node {} (d)
        edge node {} (f)
    (e) edge node {} (d)
        edge node {} (f)
    (d) edge node {} (g)
    (f) edge node {} (g)
    (h) edge node {} (i)
        edge [bend right=50] node [swap] {} (j)
    (i) edge node {} (j);
  \end{tikzpicture}
}
\hspace{10pt}
\subfloat[][\label{fig:msp}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
  \node[] (a) at (0,0) {$a$};
   \node[] (b) at (1,0) {$b$};
   \node[] (c) at (2,0.5) {$c$};
   \node[] (d) at (3,0.5) {$d$};
   \node[] (e) at (2,-0.5) {$e$};
   \node[] (f) at (3,-0.5) {$f$};
   \node[] (g) at (4,0) {$g$};
   \node[] (h) at (1,-1) {$h$};
   \node[] (i) at (2,-1) {$i$};
   \node[] (j) at (3,-1) {$j$};

  \path[->]
    (a) edge node {} (b)
        edge node {} (h)
    (b) edge node {} (c)
        edge node {} (e)
    (c) edge node {} (d)
        edge node {} (f)
    (e) edge node {} (d)
        edge node {} (f)
    (d) edge node {} (g)
    (f) edge node {} (g)
    (h) edge node {} (i)
    (i) edge node {} (j);
  \end{tikzpicture}
}
\end{center}
\caption{(a) GSP and (b) MSP DAGs}
\label{fig:gspmsp}
\end{figure}

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (S1) at (0,0) {$\mathcal{S}$};
   \node[] (P1) at (0.5,1) {$\mathcal{P}$};
   \node[] (a) at (-0.5,1) {$a$};
   \node[] (S2) at (-1,2) {$\mathcal{S}$}; %+1
   \node[] (S3) at (2,2) {$\mathcal{S}$}; %-1
   \node[] (b) at (-2,3) {$b$};
   \node[] (S4) at (0,3) {$\mathcal{S}$};
   \node[] (P2) at (-1,4) {$\mathcal{P}$};
   \node[] (g) at (1,4) {$g$};
   \node[] (S5) at (-2,5) {$\mathcal{S}$};
   \node[] (S6) at (0,5) {$\mathcal{S}$};
   \node[] (c) at (-2.5,6) {$c$};
   \node[] (d) at (-1.5,6) {$d$};
   \node[] (e) at (-0.5,6) {$e$};
   \node[] (f) at (0.5,6) {$f$};

   \node[] (h) at (1,3) {$h$};
   \node[] (S7) at (3,3) {$\mathcal{S}$};
   \node[] (i) at (2.5,4) {$i$};
   \node[] (j) at (3.5,4) {$j$};

  \path[->]
    (S1) edge node {} (a)
          edge node {} (P1)
    (P1) edge node {} (S2)
          edge node {} (S3)
    (S2) edge node {} (b)
        edge node {} (S4)
    (S4) edge node {} (P2)
          edge node {} (g)
    (P2) edge node {} (S5)
          edge node {} (S6)
    (S5) edge node {} (c)
          edge node {} (d)
    (S6) edge node {} (e)
          edge node {} (f)
    (S3) edge node {} (h)
         edge node {} (S7)
    (S7) edge node {} (i)
          edge node {} (j);
  \end{tikzpicture}
  \caption{Binary decomposition tree of the MSP of Figure~\ref{fig:msp}}
  \label{fig:t}
\end{center}
\end{figure}

Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have also proposed a linear algorithm to know if a DAG is MSP and, if it is, to decompose it to its associated binary decomposition tree. This algorithm is based on the duality of the class of MSP DAGs, with the class of \emph{Two Terminal Series Parallel} DAGs (TTSP) from which a binary tree decomposition can be performed in linear time. However, in this paper we are interested in DAGs that we already know as MSP. Thus, we only give the needed definitions to understand the binary decomposition algorithm, and not the recognition algorithm. To let the reader understand this algorithm, which is modified in this work, the needed definitions are given. 
%  Other algorithms have also been proposed, all of them in linear time~\cite{Schoenmakers95anew}.

\begin{mydef}
The \emph{line digraph} of a digraph $G$ is a digraph $L(G)$ that has:
\begin{itemize}
\item a vertex $f(e)$ for each edge $e$ of $G$; and
\item an edge $(f(e_1),f(e_2))$ for each pair of edges of $G$ of the form $e_1=(u,v)$, $e_2=(v,w)$.
\end{itemize}
\end{mydef}

\begin{mydef}Two Terminal Series Parallel
\begin{itemize}
\item A digraph consisting of two vertices joined by a single edge is TTSP.
\item If $G_1$ and $G_2$ are TTSP digraphs, so is the digraph obtained by either of the following operations:
\begin{itemize}
\item \emph{Two terminal parallel composition}: identify the sourc of $G_1$ with the source of $G_2$ and the sink of $G_1$ with the sink of $G_2$.
\item \emph{Two terminal series composition}: identify the sink of $G_1$ with the source of $G_2$.
\end{itemize}
\end{itemize}
\end{mydef}

\begin{myth}
If the DAG $G$ is a MSP graph, its \emph{inverse line DAG} $L^{-1}(G)$ is TTSP.
\end{myth}

The binary decomposition tree algorithm is based on the fact that the decomposition can be obtained as a byproduct of a reduction process. In order to obtain the decomposition, Valdes \& Al associate a label with each edge of the digraph being reduced. Initially the label of each edge is a trivial binary tree consisting of a single node. As the reduction process introduces new edges the rules of Figure~\ref{fig:rules} are used to compute the binary trees used to label them. The algorithm ends when the TTSP graph is reduced to its minimum, i.e.\ two vertices and a single edge between them.

\begin{figure}[h!]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
%FIRST RULE
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};

  \path[->]
    (x) edge node [above] {$T_1$} (y)
    (y) edge node [above] {$T_2$} (z);
  \end{tikzpicture}
  }
\hspace{50pt}
%ARROW
\subfloat[]{
$\Rightarrow$
}
\hspace{50pt}
%SECOND RULE
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
   \node[] (s) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (t1) at (1,1.5) {$T_1$};
   \node[] (t2) at (2,1.5) {$T_2$};

  \path[->]
    (x) edge node {} (y)
    (s) edge node {} (t1)
        edge node {} (t2);
  \end{tikzpicture}
}
\\
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};

  \path[->]
    (x) edge [bend left] node [above] {$T_1$} (y)
        edge node [above] {$T_2$} (y);
  \end{tikzpicture}
  }
\hspace{50pt}
\subfloat[]{
$\Rightarrow$
}
\hspace{50pt}
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
   \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
   \node[] (t1) at (1,1.5) {$T_1$};
   \node[] (t2) at (2,1.5) {$T_2$};

  \path[->]
    (x) edge node {} (y)
    (p) edge node {} (t1)
        edge node {} (t2);
  \end{tikzpicture}
}
\caption{Reduction rules of the decomposition tree algorithm.}
\label{fig:rules}
\end{center}
\end{figure}

Finally, Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identify a forbidden shape, or subgraph, called $N$ and represented in Figure~\ref{fig:n}, such that

\begin{myth}
A DAG $G$ is GSP if and only if its transitive closure does not contain $N$ as a subgraph.
\end{myth}

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (a) at (0,0) {$a$};
   \node[] (b) at (1,0) {$b$};
   \node[] (c) at (0,-1) {$c$};
   \node[] (d) at (1,-1) {$d$};

  \path[->]
    (a) edge node {} (b)
    (c) edge node {} (b)
        edge node {} (d);
  \end{tikzpicture}
  \caption{Forbidden $N$ subgraph shape for a DAG to be GSP.}
  \label{fig:n}
\end{center}
\end{figure}

%------------------
\subsection{Multi-stencil programs.}
We are interested in GSP and MSP classes of graphs to be able to represent computations of a multi-stencil program as a set of sequences and parallel executions. Actually, such a representation can directly be dumped to a parallel language, in which \emph{sequence} of instructions and \emph{parallel} execution of instructions are defined. Series-parallel trees can also be used as input of scheduling optimizations~\cite{Finta1996323,Wang20082684} to improve task parallelism efficiency, which opens more perspectives to this work. Finally, in this paper is presented that such a series-parallel representation can also be dumped to component models by defining specific control components. Thus, the proposed DSL inheritates software engineering advantages of component models, such as code re-use, productivity and maintainability.

Returning back to the parallelization formalism of Section~\ref{}, two important steps have to be performed to make $\Gamma_{hybrid}$ a MSP graph:
\begin{itemize}
\item transitive reduction,
\item deletion of the forbidden $N$ shape.
\end{itemize}

The transitive reduction has already been applied on $\Gamma_{hybrid}$ by the transitivity of the relation $\blacktriangleleft$. However, even by applying this transitive reduction it is possible to obtain a DAG which is not MSP. Actually, it is possible in a multi-stencil program to have a set of computations such that their dependencies form the forbidden $N$ subgraph. For example, if $c_0$, $c_1$, $c_2$ and $c_3$ are computations of a $\mathcal{MSP}$ such that $c_0 \prec c_1$, $c_2 \prec c_1$ and $c_2 \prec c_3$, the \emph{zigzag} relation $c_0 \prec c_1 \succ c_2 \prec c_3$ which form a forbidden subgraph of MSP DAGs is found in $\Gamma_{hybrid}$. For this reason, we have made the choice to over-constrain such a case by adding the relation $c_0 \prec c_3$ such that a complete graph is created and can be translated to a series-parallel decomposition as illustrated in Figure~\ref{fig:allover}.

\begin{figure}[h!]
\begin{center}
\subfloat[][Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$c_0$};
   \node[] (c1) at (2,0) {$c_1$};
   \node[] (c2) at (0,-1) {$c_2$};
   \node[] (c3) at (2,-1) {$c_3$};

  \path[->]
    (c0) edge node {} (c1)
          edge [dashed] node [swap] {} (c3)
    (c2) edge node {} (c1)
        edge node {} (c3);
  \end{tikzpicture}
}
\hspace{50pt}
\subfloat[][Series-parallel tree associated to the over-constraint\label{fig:treeover}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (P1) at (0,0) {$\mathcal{P}$};
   \node[] (S1) at (-1,1) {$\mathcal{S}$};
   \node[] (S2) at (1,1) {$\mathcal{S}$};
   \node[] (c0) at (-1.5,2) {$c_0$};
   \node[] (c1) at (-0.5,2) {$c_1$};
   \node[] (c2) at (0.5,2) {$c_2$};
   \node[] (c3) at (1.5,2) {$c_3$};

  \path[->]
    (P1) edge [dashed] node [swap] {} (S1)
          edge [dashed] node [swap] {} (S2)
    (S1) edge node {} (c0)
          edge node {} (c1)
    (S2) edge node {} (c2)
        edge node {} (c3);
  \end{tikzpicture}
}
\caption{Deletion of forbidden subgraphs.}
\label{fig:allover}
\end{center}
\end{figure}

This approximation on the dependencies is acceptable for multi-stencil programs, first because it rarely appends, second because of the relative homogeneity of computations. Actually, all computations, except the ones on the physical border of the domain, are performed on an entire domain of the mesh, and as a computation performs a single quantity at a time ($w_i$ of $c_i$ is a singleton), the amount of arithmetic operations in a computation are quite homogeneous.

After these over-constraints are applied, $\Gamma_{hybrid}$ is a MSP DAG. As a result, the binary tree decomposition algorithm of Valdes \& Al can be applied on $\Gamma_{hybrid}$. However, because of updated-computations inserted in $\Gamma_{data}$, we have to define additional reduction rules for the algorithm. For an updated-computation $c^*_j(c_j,\text{update}(\{w_i\text{, }\forall i\text{ }c_i \prec c_j\}))$, we define $*_j$ as the update needed by $c_j$, $*_j = \text{update}(\{w_i\text{, }\forall i\text{ }c_i \prec c_j\})$. As a result, the updated-computation $c^*_j$ is defined by $c^*_j(c_j,*_j)$. The initial label of the edges of the graph to reduce are the computations of the multi-stencil program. Four new rules are needed to perform the binary tree decomposition of $\Gamma_{hybrid}$ and are defined in Figure~\ref{fig:newrules}.

\begin{figure}[h!]
\captionsetup[subfigure]{labelformat=empty}
\begin{center}
%FIRST RULE
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};

  \path[->]
    (x) edge node [above] {$c_1^*$} (y)
    (y) edge node [above] {$c_2$} (z);
  \end{tikzpicture}
  }
\hspace{50pt}
%ARROW
\subfloat[]{
$\Rightarrow$
}
\hspace{50pt}
%TRANSFORMATION
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
   \node[] (s1) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s2) at (1,1.5) {$\mathcal{S}$};
   \node[] (star) at (0.5,2.5) {$*_1$};
   \node[] (t1) at (1.5,2.5) {$c_1$};
   \node[] (t2) at (2,1.5) {$c_2$};

  \path[->]
    (x) edge node {} (y)
    (s1) edge node {} (s2)
        edge node {} (t2)
    (s2) edge node {} (star)
         edge node {} (t1);
  \end{tikzpicture}
}
\\
%SECOND RULE
%FIRST RULE
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (z) at (3,0) {};

  \path[->]
    (x) edge node [above] {$c_1^*$} (y)
    (y) edge node [above] {$c_2^*$} (z);
  \end{tikzpicture}
  }
\hspace{50pt}
%ARROW
\subfloat[]{
$\Rightarrow$
}
\hspace{50pt}
%TRANSFORMATION
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
   \node[] (s1) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s2) at (0.5,1.5) {$\mathcal{S}$};
   \node[] (s3) at (2.5,1.5) {$\mathcal{S}$};
   \node[] (star1) at (0,2.5) {$*_1$};
   \node[] (t1) at (1,2.5) {$c_1$};
   \node[] (star2) at (2,2.5) {$*_2$};
   \node[] (t2) at (3,2.5) {$c_2$};

  \path[->]
    (x) edge node {} (y)
    (s1) edge node {} (s2)
        edge node {} (s3)
    (s2) edge node {} (star1)
         edge node {} (t1)
    (s3) edge node {} (star2)
         edge node {} (t2);
  \end{tikzpicture}
}
\\
%THIRD RULE
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};

  \path[->]
    (x) edge [bend left] node [above] {$c_1^*$} (y)
        edge [bend right] node [below] {$c_2^*$} (y);
  \end{tikzpicture}
  }
\hspace{50pt}
%ARROW
\subfloat[]{
$\Rightarrow$
}
\hspace{50pt}
%TRANSFORMATION
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
   \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
   \node[] (s) at (1,1.5) {$\mathcal{S}$};
   \node[] (star) at (0.5,2.5) {$*_1$};
   \node[] (t1) at (1.5,2.5) {$c_1$};
   \node[] (t2) at (2,1.5) {$c_2$};

  \path[->]
    (x) edge node {} (y)
    (p) edge node {} (s)
        edge node {} (t2)
    (s) edge node {} (star)
        edge node {} (t1);
  \end{tikzpicture}
}
\\
%FOURTH RULE
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (1.5,0) {};

  \path[->]
    (x) edge [bend left] node [above] {$c_1^*$} (y)
        edge [bend right] node [below] {$c_2^*$} (y);
  \end{tikzpicture}
  }
\hspace{50pt}
%ARROW
\subfloat[]{
$\Rightarrow$
}
\hspace{50pt}
%TRANSFORMATION
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (x) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (y) at (3,0) {};
   \node[] (p) at (1.5,0.5) {$\mathcal{P}$};
   \node[] (s2) at (0.5,1.5) {$\mathcal{S}$};
   \node[] (s3) at (2.5,1.5) {$\mathcal{S}$};
   \node[] (star1) at (0,2.5) {$*_1$};
   \node[] (t1) at (1,2.5) {$c_1$};
   \node[] (star2) at (2,2.5) {$*_2$};
   \node[] (t2) at (3,2.5) {$c_2$};

  \path[->]
    (x) edge node {} (y)
    (p) edge node {} (s2)
        edge node {} (s3)
    (s2) edge node {} (star1)
         edge node {} (t1)
    (s3) edge node {} (star2)
         edge node {} (t2);
  \end{tikzpicture}
}
\caption{Reduction rules of the decomposition tree algorithm.}
\label{fig:newrules}
\end{center}
\end{figure}

Using those additional reduction rules, the binary series-parallel tree decomposition can be computed in linear time. A complete example of reduction is given in Figure~\ref{fig:example}. The initial $\Gamma_{hybrid}$ DAG and its inverse line graph are given first, followed by the set of reductions. The label of the final single edge corresponds to the binary tree decomposition of the initial $\Gamma_{hybrid}$ graph.

%====BIG EXAMPLE
\begin{figure}[ht!]
\begin{center}
% INITIAL GRAPH
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$c_0$};
   \node[] (c1) at (1,0) {$c_1^*$};
   \node[] (c2) at (2,0.5) {$c_2$};
   \node[] (c3) at (2,-0.5) {$c_3$};
   \node[] (c4) at (3,0.5) {$c_4^*$};
   \node[] (c5) at (3,-0.5) {$c_5$};
   \node[] (c6) at (4,0.5) {$c_6$};
   \node[] (c7) at (5,0) {$c_7$};
   \node[] (c8) at (6,0) {$c_8^*$};

  \path[->]
    (c0) edge node {} (c1)
    (c1) edge node {} (c2)
         edge node {} (c3)
    (c2) edge node {} (c4)
    (c4) edge node {} (c6)
    (c3) edge node {} (c5)
    (c5) edge node {} (c7)
    (c6) edge node {} (c7)
    (c7) edge node {} (c8);
  \end{tikzpicture}
}
\\
% INVERSE LINE GRAPH
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c0d) at (1,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (3,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (3,-0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c7d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};

  \path[->]
    (c0s) edge node [above] {$c_0$} (c0d)
    (c0d) edge node [above] {$c_1^*$} (c1d)
    (c1d) edge node [above] {$c_2$} (c2d)
          edge node [above] {$c_3$} (c3d)
    (c2d) edge node [above] {$c_4^*$} (c4d)
    (c4d) edge node [above] {$c_6$} (c6d)
    (c6d) edge node [above] {$c_7$} (c7d)
    (c7d) edge node [above] {$c_8$} (c8d)
    (c3d) edge node [above] {$c_5$} (c6d);
  \end{tikzpicture}
}
\\
% TRANSFORMATIONS
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (3,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (3,-0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};

   %reduction c0 c1
   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,1.5) {$c_0$};
   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
   \node[] (star1) at (1,2.5) {$*_1$};
   \node[] (c1) at (2,2.5) {$c_1$};

   %reduction c7 c8
   \node[] (s3) at (6,0.5) {$\mathcal{S}$};
   \node[] (c7) at (5.5,1.5) {$c_7$};
   \node[] (s4) at (6.5,1.5) {$\mathcal{S}$};
   \node[] (star8) at (6,2.5) {$*_8$};
   \node[] (c8) at (7,2.5) {$c_8$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge node [above] {$c_2$} (c2d)
          edge node [above] {$c_3$} (c3d)
    (c2d) edge node [above] {$c_4^*$} (c4d)
    (c4d) edge node [above] {$c_6$} (c6d)
    (c6d) edge node [above] {} (c8d)
    (c3d) edge node [above] {$c_5$} (c6d)

    %reduction c0 c1
    (s1) edge node {} (c0)
         edge node {} (s2)
    (s2) edge node {} (star1)
         edge node {} (c1)
    %reduction c7 c8
    (s3) edge node {} (c7)
         edge node {} (s4)
    (s4) edge node {} (star8)
         edge node {} (c8);
  \end{tikzpicture}
}
\\
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};

   %reduction c0 c1
   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,1.5) {$c_0$};
   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
   \node[] (star1) at (1,2.5) {$*_1$};
   \node[] (c1) at (2,2.5) {$c_1$};

   %reduction c7 c8
   \node[] (s3) at (6,0.5) {$\mathcal{S}$};
   \node[] (c7) at (5.5,1.5) {$c_7$};
   \node[] (s4) at (6.5,1.5) {$\mathcal{S}$};
   \node[] (star8) at (6,2.5) {$*_8$};
   \node[] (c8) at (7,2.5) {$c_8$};

   %reduction c3 c5
   \node[] (s5) at (3.5,-2) {$\mathcal{S}$};
   \node[] (c3) at (3,-1) {$c_3$};
   \node[] (c5) at (4,-1) {$c_5$};

   %reduction c2 c4
   \node[] (s6) at (3.5,1) {$\mathcal{S}$};
   \node[] (c2) at (3,2) {$c_2$};
   \node[] (s7) at (4,2) {$\mathcal{S}$};
   \node[] (star4) at (3.5,3) {$*_4$};
   \node[] (c4) at (4.5,3) {$c_4$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge [bend left] node [above] {} (c4d)
    (c4d) edge node [above] {$c_6$} (c6d)
    (c6d) edge node [above] {} (c8d)
    (c1d) edge [bend right] node [above] {} (c6d)

    %reduction c0 c1
    (s1) edge node {} (c0)
         edge node {} (s2)
    (s2) edge node {} (star1)
         edge node {} (c1)
    %reduction c7 c8
    (s3) edge node {} (c7)
         edge node {} (s4)
    (s4) edge node {} (star8)
         edge node {} (c8)

    %reduction c3 c5
    (s5) edge node {} (c3)
         edge node {} (c5)

    %reduction c7 c8
    (s6) edge node {} (c2)
         edge node {} (s7)
    (s7) edge node {} (star4)
         edge node {} (c4);
  \end{tikzpicture}
}
\end{center}
\caption{Big example}
\label{fig:example}
\end{figure}

\begin{figure}[ht!]
\ContinuedFloat
\begin{center}
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (5,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (7,0) {};

   %reduction c0 c1
   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,1.5) {$c_0$};
   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
   \node[] (star1) at (1,2.5) {$*_1$};
   \node[] (c1) at (2,2.5) {$c_1$};

   %reduction c7 c8
   \node[] (s3) at (6,0.5) {$\mathcal{S}$};
   \node[] (c7) at (5.5,1.5) {$c_7$};
   \node[] (s4) at (6.5,1.5) {$\mathcal{S}$};
   \node[] (star8) at (6,2.5) {$*_8$};
   \node[] (c8) at (7,2.5) {$c_8$};

   %reduction c3 c5
   \node[] (s5) at (3.5,-2) {$\mathcal{S}$};
   \node[] (c3) at (3,-1) {$c_3$};
   \node[] (c5) at (4,-1) {$c_5$};

   %reduction c2 c6
   \node[] (s8) at (3.5,1) {$\mathcal{S}$};
   \node[] (c6) at (4,2) {$c_6$};
   \node[] (s6) at (3,2) {$\mathcal{S}$};
   \node[] (c2) at (2.5,3) {$c_2$};
   \node[] (s7) at (3.5,3) {$\mathcal{S}$};
   \node[] (star4) at (3,4) {$*_4$};
   \node[] (c4) at (4,4) {$c_4$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge [bend left] node [above] {} (c6d)
    (c6d) edge node [above] {} (c8d)
    (c1d) edge [bend right] node [above] {} (c6d)

    %reduction c0 c1
    (s1) edge node {} (c0)
         edge node {} (s2)
    (s2) edge node {} (star1)
         edge node {} (c1)
    %reduction c7 c8
    (s3) edge node {} (c7)
         edge node {} (s4)
    (s4) edge node {} (star8)
         edge node {} (c8)

    %reduction c3 c5
    (s5) edge node {} (c3)
         edge node {} (c5)

    %reduction c2 c6
    (s8) edge node {} (s6)
         edge node {} (c6)
    (s6) edge node {} (c2)
         edge node {} (s7)
    (s7) edge node {} (star4)
         edge node {} (c4);
  \end{tikzpicture}
}
\\
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (2,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (8,0) {};

   %reduction c0 c1
   \node[] (s1) at (1,0.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,1.5) {$c_0$};
   \node[] (s2) at (1.5,1.5) {$\mathcal{S}$};
   \node[] (star1) at (1,2.5) {$*_1$};
   \node[] (c1) at (2,2.5) {$c_1$};

   %reduction c7 c8
   \node[] (s3) at (7,0.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,1.5) {$c_7$};
   \node[] (s4) at (7.5,1.5) {$\mathcal{S}$};
   \node[] (star8) at (7,2.5) {$*_8$};
   \node[] (c8) at (8,2.5) {$c_8$};

   %reduction c2 c4 c6 c3 c5
   \node[] (p1) at (4,0.5) {$\mathcal{P}$};
   \node[] (s5) at (3,1) {$\mathcal{S}$};
   \node[] (c3) at (2.5,2) {$c_3$};
   \node[] (c5) at (3.5,2) {$c_5$};
   \node[] (s8) at (5,1) {$\mathcal{S}$};
   \node[] (c6) at (5.5,2) {$c_6$};
   \node[] (s6) at (4.5,2) {$\mathcal{S}$};
   \node[] (c2) at (4,3) {$c_2$};
   \node[] (s7) at (5,3) {$\mathcal{S}$};
   \node[] (star4) at (4.5,4) {$*_4$};
   \node[] (c4) at (5.5,4) {$c_4$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge node [above] {} (c6d)
    (c6d) edge node [above] {} (c8d)

    %reduction c0 c1
    (s1) edge node {} (c0)
         edge node {} (s2)
    (s2) edge node {} (star1)
         edge node {} (c1)
    %reduction c7 c8
    (s3) edge node {} (c7)
         edge node {} (s4)
    (s4) edge node {} (star8)
         edge node {} (c8)

    %reduction c2 c4 c6 c3 c5
    (p1) edge node {} (s5)
         edge node {} (s8)
    (s5) edge node {} (c3)
         edge node {} (c5)
    (s8) edge node {} (s6)
         edge node {} (c6)
    (s6) edge node {} (c2)
         edge node {} (s7)
    (s7) edge node {} (star4)
         edge node {} (c4);
  \end{tikzpicture}
}
\\
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (8,0) {};

   %all
   \node[] (s9) at (2.5,1.5) {$\mathcal{S}$};
   \node[] (s10) at (4,0.5) {$\mathcal{S}$};

   %reduction c0 c1
   \node[] (s1) at (1,2) {$\mathcal{S}$};
   \node[] (c0) at (0.5,3) {$c_0$};
   \node[] (s2) at (1.5,3) {$\mathcal{S}$};
   \node[] (star1) at (1,4) {$*_1$};
   \node[] (c1) at (2,4) {$c_1$};

   %reduction c7 c8
   \node[] (s3) at (7,1) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2) {$c_7$};
   \node[] (s4) at (7.5,2) {$\mathcal{S}$};
   \node[] (star8) at (7,3) {$*_8$};
   \node[] (c8) at (8,3) {$c_8$};

   %reduction c2 c4 c6 c3 c5
   \node[] (p1) at (4,2) {$\mathcal{P}$};
   \node[] (s5) at (3,2.5) {$\mathcal{S}$};
   \node[] (c3) at (2.5,3.5) {$c_3$};
   \node[] (c5) at (3.5,3.5) {$c_5$};
   \node[] (s8) at (5,2.5) {$\mathcal{S}$};
   \node[] (c6) at (5.5,3.5) {$c_6$};
   \node[] (s6) at (4.5,3.5) {$\mathcal{S}$};
   \node[] (c2) at (4,4.5) {$c_2$};
   \node[] (s7) at (5,4.5) {$\mathcal{S}$};
   \node[] (star4) at (4.5,5.5) {$*_4$};
   \node[] (c4) at (5.5,5.5) {$c_4$};

  \path[->]
    (c0s) edge node {} (c8d)

    %reduction c0 c1
    (s1) edge node {} (c0)
         edge node {} (s2)
    (s2) edge node {} (star1)
         edge node {} (c1)
    %reduction c7 c8
    (s3) edge node {} (c7)
         edge node {} (s4)
    (s4) edge node {} (star8)
         edge node {} (c8)

    %reduction c2 c4 c6 c3 c5
    (p1) edge node {} (s5)
         edge node {} (s8)
    (s5) edge node {} (c3)
         edge node {} (c5)
    (s8) edge node {} (s6)
         edge node {} (c6)
    (s6) edge node {} (c2)
         edge node {} (s7)
    (s7) edge node {} (star4)
         edge node {} (c4)

    %all
    (s9) edge node {} (s1)
         edge node {} (p1)
    (s10) edge node {} (s9)
          edge node {} (s3);
  \end{tikzpicture}
}
\end{center}
\caption{Big example}
\label{fig:example}
\end{figure}

At this point, the $\mathcal{MSP}$ program can be dumped to any parallel language in which a sequence and a parallel execution are available. For example, we could imagine a parallel functionnal language with a binary \emph{sequence} function and a binary \emph{parallel} function such that the tree of Figure~\ref{fig:t} corresponds the following nested call:

\begin{equation}
\begin{split}
\text{sequence}(a,\text{parallel}(&\\
&\text{sequence}(b,\text{sequence}(\text{parallel}(\text{sequence}(c,d),\text{sequence}(e,f)),g)),\\
&\text{sequence}(h,\text{sequence}(i,j))\\
&))
\end{split}
\end{equation}