### 2015-07-31 WOLFHPC15/article/component.tex
fix all helene 30_07

%----------------------------------------
%\subsection{Overview}
%----------------------------------------
This section describes both the projection of $\Gamma_{tsp}$ to
components, and the remaining needed components to execute a
multi-stencil application. Before those details the section gives an overview of
component models, and introduce the concepts of the Low Level Component Model \llc, which is used in this paper.

%----------------------------------------
\subsection{Component Model Overview}
%----------------------------------------
Component model is a domain of software
engineering~\cite{Szyperski:2002:CSB:515228}, which promotes code
re-use, separations of concerns, and thus maintainability. An
application is made of a set of components; a component being a kind of
black box representing an independent functionnality of the application, and which interacts only through its ports. A port specifies the services provided and required by the component.
%
With respect to high performance computing, some works have also shown
that component models can achieve the needed level of performance, and
scalability while also helping in application
portability~\cite{Bernholdt01052006, bigot:inria-00388508, UCHPC2015}

Many component models exist, each of them with its own specifications
and singularities. In distributed computing, well known component
models are CCM~\cite{corba:omg06} (CORBA Component Model), and
GCM~\cite{Baude} (Grid Component Model); in HPC, there are
CCA~\cite{Bernholdt01052006} (Common Component Architecture), and
\llc~\cite{l2c} (Low Level Component), for example.
%
As this work makes use of \llc, in particular for the experiments, let
introduce concepts of this component model with more details.

%----------------------------------------
\subsection{\llc}
%----------------------------------------

\llc is a minimalist \texttt{C++} based HPC-oriented component model
where a component extends the concept of class by specifying in its
interfaces the services that it offers ($provide$ ports) and that it
needs, either a single service instance ($use$ ports), or multiple
service instances ($use-multiple$ ports). Services are \texttt{C++}
interfaces. \llc also offers $MPI$ ports that enable components to
share MPI communicators. Components can also have attribute ports to
be configured.
%
As illustrated in Figures~\ref{fig:ports}, a $provide$ port is
represented with a white circle, a $use$
port with a black circle, a $use-multiple$ port by a black circle with
a white $m$ in it. MPI port are
connected with a black rectangle.

A classical \llc-based application is a static \emph{assembly} of
components made of instances of components and of connections between
component ports. Such an assembly is described in LAD, an XML
dialect~\cite{l2c}.

\begin{figure}[t]
\begin{center}
\subfloat[][\label{fig:2comp}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$C_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[provide,right=1.5cm of u] (p1) {};
   \node[component,right=1.5cm of p1] (C1) {$C_1$};
   \node[use,right=1.5cm of C1] (um) {$m$};

  \path[-]
    (p) edge node {$p$} (C)
    (C.east) edge node {$u$} (u)
    (C1)  edge node {$v$} (um)
    (p1) edge node {$q$} (C1);
\end{tikzpicture}
}
\subfloat[][\label{fig:ass}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$c_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[component,right=1.5 of u] (C1) {$c_1$};
   \node[use,right=1.5 of C1] (um) {$m$};

  \path[-]
    (p) edge node {$p$} (C)
    (C) edge node {$u$} (u)
    (C1)  edge node {$v$} (um)
    (u) edge node {$q$} (C1);
\end{tikzpicture}
}
\subfloat[][\label{fig:mpi}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$c_2$};
   \node[provide] (p) at (-1.5,0) {};
   \node[mpi] (u) at (1.5,0) {};
   \node[component,right=1.5 of u] (C1) {$c_3$};

  \path[-]
    (p) edge node {$r$} (C)
    (C) edge node {$m_1$} (u)
    (u) edge node {$m_2$} (C1);
\end{tikzpicture}
}
\caption{Example of components and their ports representation. a) Component $c_0$ has a provide port ($p$) and a use port ($u$); Component $c_1$ has also a provide port ($q$) but also a use multiple port ($v$). b) A use port is connected to a (compatible) provide port. c) Component $c_2$ and $c_3$ shares an MPI communicator.}
\label{fig:ports}
\end{figure}

%% In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}. As the use port is on the right of its component, and the provide port on the left, in a component assembly the component on the left uses the provide port of the component on the right.

%----------------------------------------
\subsection{MSCAC Runtime Overview}
%----------------------------------------

Figure~\ref{fig:mscac:assembly} gives an overview of the back-end component
assembly that is executed. As discusses in Section~\ref{sect:mscac},
there are different types of components. Components~\texttt{Driver},
\texttt{DriverApp}, and \texttt{Time} are always instanciated in the final component assembly; the input MSL file is not used by them except the component \texttt{Time} which uses the terminal \texttt{iteration} of the grammar. There is also a
single Component~\texttt{DDS} (the current version handles a single mesh
type). All these components are provided with the
compiler; dumping this part of the assembly is straightforward.

While Component~\texttt{DDS} is managing the structure of the mesh,
each data of the simulation is handled by a component of type
\texttt{Data}. Therefore, the compiler generates in the assembly as
many instances of such component as needed, from the \texttt{data}
section of an MSL program.

The last part of the assembly is the computation part. Each computation kernel is embedded in a kernel component, denoted K, which have to be written by the user and provided
to the compiler. A kernel component encapsulates a computations. To
access data, a kernel component is connected, by the compiler, to the adequate
\texttt{Data} components such that their ports name respect the names
given in the \texttt{data} section of MSL. We denote the fact that they may have
several use port by adding a star on the use port, as illustrated in Figure~\ref{fig:k}.
\begin{figure}
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (k) at (0,0) {$K$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[right=0.2 of u] (star) {*};

  \path[-]
    (p) edge node {} (k)
    (k) edge node {} (u);
\end{tikzpicture}
\end{center}
\caption{Kernel component.}
\label{fig:k}
\end{figure}

%% In the computation part of the simulation a final component is needed,
%% the computation component, denoted K. This type of component
%% represents a single computation of the simulation, i.e. the
%% computation of a single output data (written) using a set of input
%% data (read). Using a \emph{use-multiple}, however, a list of component
%% is used and the explicit identification of each component to use is
%% lost. If for SEQ or PAR, the identification of components is not
%% usefull, it is usefull to manipulate data in a computation. For this
%% reason a computation component has as much \emph{use} ports as input
%% and output data to manipulate. We denote this in the bellow
%% representation as a star on the use port.

The rest of the computation part of the assembly contains the static parallel schedule
computed by MSC, \ie $\Gamma_{tsp}$. This is achieved through the use of
three specific components which manage the $P$, $S$, and $sync$
operations of $\Gamma_{tsp}$ as explained hereafter.

%----------------------------------------
\subsection{Control Components}
%----------------------------------------
The series-parallel tree decomposition $\Gamma_{tsp}$ represents the
control of the dependencies of the simulation. As a result to be able
to dump it to a component assembly it is needed to introduce what can
be called \emph{control components}. Those components can be used for
any case of simulation which increases code reuse between
simulations. A control component is composed of a single
\emph{provide} port linked to a single execution method, most of the
time called the \emph{go} method. We introduce three types of control
components represented in Figure~\ref{fig:ctrlcomponents}.

\begin{figure}[t]
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (seq) at (0,0) {$SEQ$}; \node[provide] (p) at
   (-1,0) {}; \node[use] (u) at (1,0) {$m$};

  \path[-]
    (p) edge node {} (seq)
    (seq) edge node {} (u);
\end{tikzpicture}
\hspace{\fill}
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (seq) at (0,0) {$PAR$};
   \node[provide] (p) at (-1,0) {};
   \node[use] (u) at (1,0) {$m$};

  \path[-]
    (p) edge node {} (seq)
    (seq) edge node {} (u);
\end{tikzpicture}
\hspace{\fill}
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (sync) at (0,0) {$SYNC$};
   \node[provide] (p) at (-1,0) {};
   \node[use] (u) at (1,0) {};

  \path[-]
    (p) edge node {} (sync)
    (sync) edge node {} (u);
\end{tikzpicture}
\\
a) Comp. SEQ
\hspace{\fill}
b) Comp. PAR
\hspace{\fill}
c) Comp. SYNC
\\
\caption{The three control components used by the back-end.}
\label{fig:ctrlcomponents}
\end{figure}

\begin{description}
\item[Sequence component (SEQ)] It is the direct representation of a sequence node of $\Gamma_{tsp}$. The role of this component is to call an ordered list of other components. Its interface contains an ordered \emph{use-multiple} port to be connected to the components to call in sequence.

\item[Parallel component (PAR)] It is the direct representation of a parallel node of $\Gamma_{tsp}$. The role of this component is to call simultaneously a set of other components. It offers a \emph{use-multiple} port to be connected to the components to call in parallel.

\item[Synchronization component (SYNC)]. It is the direct representation to an update leaf of $\Gamma_{tsp}$. The role of this component is to call the synchronization of a given data. It offers a \emph{use} port to be connected to the data to update.

\end{description}

%----------------------------------------
\subsection{Dump To Component Assembly}
%----------------------------------------
From $\Gamma_{tsp}$ and using the control components described above,
a direct dump can be done to a component assembly for an hybrid (data
and task) parallelization of a simulation. 
\fix{HC: ce qui suit est a changer avec la fusion du code de plusieurs composants}
However, it is also
possible to dump to a data parallelization only. In such a case the
computation of $\Gamma_{sync}$ is enough to generate the component
assembly. Actually a single SEQ component is thus needed and this
component is linked to all computations and synchronizations of
$\Gamma_{sync}$ directly.

Figure~\ref{fig:tsp:assembly} displays the assembly part corresponding to
$\Gamma_{tsp}$ of Figure~\ref{fig:tsp}. In this figure, the ports
linked to data (use and use-multiple ports of SYNC and K) are
represented but are not connected. Moreover, each computation
component is an instance of the component K presented before but using
the identification name of the computation in the MSL file.

\begin{figure*}[t]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   %seq0
   \node[component] (seq0) at (0,0) {$SEQ$};
   \node[provide] (seq0p) [left = 0.8cm of seq0] {};
   \node[use] (seq0u) [right = 1cm of seq0] {$m$};
   %c0
   \node[component] (c0) [right = 2cm of seq0u] {$c_0$};
   \node[use] (c0u) [right = 0.8cm of c0] {};
   \node[right=0.2 of c0u] (star) {*};
   %sync0
   \node[component] (sync0) [below = 1cm of c0] {$SYNC$};
   \node[use] (sync0u) [right = 1cm of sync0] {$m$};
   %c1
   \node[component] (c1) [below = 0.8cm of sync0] {$c_1^*$};
   \node[use] (c1u) [right = 0.8cm of c1] {};
   \node[right=0.2 of c1u] (star) {*};
   %par0
   \node[component] (par0) [below = 0.8cm of c1] {$PAR$};
   \node[use] (par0u) [right = 1cm of par0] {$m$};
  %seq1
   \node[component] (seq1) [right = 1cm of par0u] {$SEQ$};
   \node[use] (seq1u) [right = 1cm of seq1] {$m$};
   %par1
   \node[component] (par1) [right = 1cm of seq1u] {$PAR$};
   \node[use] (par1u) [right = 1cm of par1] {$m$};
  %c2
   \node[component] (c2) [right = 1cm of par1u] {$c_2$};
   \node[use] (c2u) [right = 0.8cm of c2] {};
   \node[right=0.2 of c2u] (star) {*};
  %sync1
   \node[component] (sync1) [below = 0.8cm of c2] {$SYNC$};
   \node[use] (sync1u) [right = 1cm of sync1] {$m$};
  %c4
   \node[component] (c4) [below = 0.8cm of par1] {$c_4^*$};
   \node[use] (c4u) [right = 0.8cm of c4] {};
   \node[right=0.2 of c4u] (star) {*};
  %c6
   \node[component] (c6) [below = 0.8cm of c4] {$c_6$};
   \node[use] (c6u) [right = 0.8cm of c6] {};
   \node[right=0.2 of c6u] (star) {*};
  %seq2
   \node[component] (seq2) [below = 2.5cm of seq1] {$SEQ$};
   \node[use] (seq2u) [right = 1cm of seq2] {$m$};
  %c3
   \node[component] (c3) [right = 1cm of seq2u] {$c_3$};
   \node[use] (c3u) [right = 0.8cm of c3] {};
   \node[right=0.2 of c3u] (star) {*};
  %c5
   \node[component] (c5) [below = 0.8cm of c3] {$c_5$};
   \node[use] (c5u) [right = 0.8cm of c5] {};
   \node[right=0.2 of c5u] (star) {*};
  %c7
   \node[component] (c7) [below = 0.8cm of par0] {$c_7$};
   \node[use] (c7u) [right = 0.8cm of c7] {};
   \node[right=0.2 of c7u] (star) {*};
  %sync2
   \node[component] (sync2) [below = 0.8cm of c7] {$SYNC$};
   \node[use] (sync2u) [right = 1cm of sync2] {$m$};
  %c8
   \node[component] (c8) [below = 0.8cm of sync2] {$c_8^*$};
   \node[use] (c8u) [right = 0.8cm of c8] {};
   \node[right=0.2 of c8u] (star) {*};

   \path[-]
   %seq0
    (seq0) edge node {} (seq0u)
    (seq0p) edge node {} (seq0)
   %c0
    (seq0u) edge node {} (c0)
    (c0) edge node {} (c0u)
    %sync0
    (seq0u) edge node {} (sync0.west)
    (sync0) edge node {} (sync0u)
    %c1
    (seq0u) edge node {} (c1)
    (c1) edge node {} (c1u)
   %par0
    (seq0u) edge node {} (par0.west)
    (par0) edge node {} (par0u)
  %seq1
    (par0u) edge node {} (seq1)
    (seq1) edge node {} (seq1u)
  %par1
    (seq1u) edge node {} (par1.west)
    (par1) edge node {} (par1u)
  %c2
    (par1u) edge node {} (c2)
    (c2) edge node {} (c2u)
  %sync0
    (par1u) edge node {} (sync1.west)
    (sync1) edge node {} (sync1u)
  %c4
    (seq1u) edge node {} (c4.west)
    (c4) edge node {} (c4u)
  %c6
    (seq1u) edge node {} (c6.west)
    (c6) edge node {} (c6u)
  %seq2
    (par0u) edge node {} (seq2.west)
    (seq2) edge node {} (seq2u)
  %c3
    (seq2u) edge node {} (c3)
    (c3) edge node {} (c3u)
  %c6
    (seq2u) edge node {} (c5.west)
    (c5) edge node {} (c5u)
  %c6
    (seq0u) edge node {} (c7.west)
    (c7) edge node {} (c7u)
  %sync2
    (seq0u) edge node {} (sync2.west)
    (sync2) edge node {} (sync2u)
  %c8
    (seq0u) edge [bend right] node {} (c8.west)
    (c8) edge node {} (c8u)
   ;
\end{tikzpicture}
\caption{Component assembly representing the computation part generated from $\Gamma_{tsp}$ of Figure~\ref{fig:tsp}.}
\label{fig:tsp:assembly}
\end{center}
\end{figure*}

%----------------------------------------