### 2015-08-03 WOLFHPC15/article/component.tex
5.4,5.5

%----------------------------------------
%\subsection{Overview}
%----------------------------------------
This section describes the projection of a MSP program to components, including the components used to implement the static scheduling of $\Gamma_{tsp}$.
Before those details the section gives an overview of component models, and introduce the Low Level Components (\llc), the model used in this paper.

%----------------------------------------
\subsection{Component-Based Software Engineering}
%----------------------------------------
Component-based software engineering (CBSE) is a domain of software engineering~\cite{Szyperski:2002:CSB:515228} which promotes code
re-use, separations of concerns, and thus maintainability.
An application is made of a set of component instances.
A component is a black box that implements an independent functionality of the application, and which interacts with its environment only through well defined interfaces: its ports.
A port can for example specify services provided or required by the component.
With respect to high performance computing, some works have also shown
that component models can achieve the needed level of performance, and
scalability while also helping in application
portability~\cite{Bernholdt01052006, bigot:inria-00388508, UCHPC2015}

Many component models exist, each of them with its own specificities.
Well known component models include for example the CORBA Component Model (CCM)~\cite{corba:omg06}, and the Grid Component Model (GCM)~\cite{Baude} in distributed computing, the Common Component Architecture (CCA)~\cite{Bernholdt01052006}, and Low Level Components (\llc)~\cite{l2c} in HPC.
This work makes use of \llc for the experiments.
Let us therefore introduce the concepts of this model in more details.

%----------------------------------------
\subsection{\llc}
%----------------------------------------

\llc is a minimalist \texttt{C++} based HPC-oriented component model
where a component extends the concept of class.
The services offered by the components are specified trough $provide$ ports,
those used either by $use$ ports for a single service instance,
or $use-multiple$ ports for multiple service instances.
Services are specified as \texttt{C++} interfaces.
\llc also offers $MPI$ ports that enable components to share MPI communicators.
Components can also have attribute ports to be configured.
%
As illustrated in Figures~\ref{fig:ports}, a $provide$ port is
represented by a white circle, a $use$
port with a black circle, a $use-multiple$ port by a black circle with
a white $m$ in it. MPI port are
connected with a black rectangle.

\todo[inline]{JB: chez moi la figure\,\ref{fig:ports} deborde...}

A \llc-based application is a static \emph{assembly} of components instances and connections between their ports.
Such an assembly is described in LAD, an XML dialect~\cite{l2c}.

\begin{figure}[t]
\begin{center}
\subfloat[][\label{fig:2comp}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$C_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[provide,right=1.5cm of u] (p1) {};
   \node[component,right=1.5cm of p1] (C1) {$C_1$};
   \node[use,right=1.5cm of C1] (um) {$m$};

  \path[-]
    (p) edge node {$p$} (C)
    (C.east) edge node {$u$} (u)
    (C1)  edge node {$v$} (um)
    (p1) edge node {$q$} (C1);
\end{tikzpicture}
}
\subfloat[][\label{fig:ass}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$c_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[component,right=1.5 of u] (C1) {$c_1$};
   \node[use,right=1.5 of C1] (um) {$m$};

  \path[-]
    (p) edge node {$p$} (C)
    (C) edge node {$u$} (u)
    (C1)  edge node {$v$} (um)
    (u) edge node {$q$} (C1);
\end{tikzpicture}
}
\subfloat[][\label{fig:mpi}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$c_2$};
   \node[provide] (p) at (-1.5,0) {};
   \node[mpi] (u) at (1.5,0) {};
   \node[component,right=1.5 of u] (C1) {$c_3$};

  \path[-]
    (p) edge node {$r$} (C)
    (C) edge node {$m_1$} (u)
    (u) edge node {$m_2$} (C1);
\end{tikzpicture}
}
\caption{Example of components and their ports representation. a) Component $c_0$ has a provide port ($p$) and a use port ($u$); Component $c_1$ has also a provide port ($q$) but also a use multiple port ($v$). b) A use port is connected to a (compatible) provide port. c) Component $c_2$ and $c_3$ shares an MPI communicator.}
\label{fig:ports}
\end{center}
\end{figure}

%% In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}. As the use port is on the right of its component, and the provide port on the left, in a component assembly the component on the left uses the provide port of the component on the right.

%----------------------------------------
\subsection{MSCAC Runtime Overview}
%----------------------------------------

Figure~\ref{fig:mscac:assembly} gives an overview of the back-end component
assembly that is executed. As discusses in Section~\ref{sect:mscac},
there are different types of components. Components~\texttt{Driver},
\texttt{DriverApp}, and \texttt{Time} are always instanciated in the final component assembly; the input MSL file is not used by them except the component \texttt{Time} which uses the terminal \texttt{iteration} of the grammar. There is also a
single Component~\texttt{DDS} (the current version handles a single mesh
type). All these components are provided with the
compiler; dumping this part of the assembly is straightforward.

While Component~\texttt{DDS} is managing the structure of the mesh,
each data of the simulation is handled by a component of type
\texttt{Data}. Therefore, the compiler generates in the assembly as
many instances of such component as needed, from the \texttt{data}
section of an MSL program.

The last part of the assembly is the computation part.
Each computation kernel is embedded in a different kernel component which has to be written by the user and provided
to the compiler.
We denote this kind of components K.
A kernel component encapsulates a computations. To
access data, a kernel component is connected, by the compiler, to the adequate
\texttt{Data} components such that their ports name respect the names
given in the \texttt{data} section of MSL. We denote the fact that they may have
several use port by adding a star on the use port, as illustrated in Figure~\ref{fig:k}.
\begin{figure}
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (k) at (0,0) {$K$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[right=0.2 of u] (star) {*};

  \path[-]
    (p) edge node {} (k)
    (k) edge node {} (u);
\end{tikzpicture}
\end{center}
\caption{Kernel component.}
\label{fig:k}
\end{figure}

%% In the computation part of the simulation a final component is needed,
%% the computation component, denoted K. This type of component
%% represents a single computation of the simulation, i.e. the
%% computation of a single output data (written) using a set of input
%% data (read). Using a \emph{use-multiple}, however, a list of component
%% is used and the explicit identification of each component to use is
%% lost. If for SEQ or PAR, the identification of components is not
%% usefull, it is usefull to manipulate data in a computation. For this
%% reason a computation component has as much \emph{use} ports as input
%% and output data to manipulate. We denote this in the bellow
%% representation as a star on the use port.

The rest of the computation part of the assembly contains the static parallel schedule
computed by MSC, \ie $\Gamma_{tsp}$. This is achieved through the use of
three specific components which manage the $P$, $S$, and $sync$
operations of $\Gamma_{tsp}$ as explained hereafter.

%----------------------------------------
\subsection{Control Components}
%----------------------------------------
The series-parallel tree decomposition $\Gamma_{tsp}$ represents the static scheduling of the simulation.
It is dumped to components by generating an assembly that exactly matches the tree.
We introduce what we call \emph{control components} to represent the nodes of the tree.
These components can be used for any simulation, which increases code reuse between simulations.
A control component exposes a single \emph{provide} port containing a single method, typically a control-only method without any parameter.
We introduce three control components represented in Figure~\ref{fig:ctrlcomponents}.

\begin{figure}[t]
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (seq) at (0,0) {$SEQ$}; \node[provide] (p) at
   (-1,0) {}; \node[use] (u) at (1,0) {$m$};

  \path[-]
    (p) edge node {} (seq)
    (seq) edge node {} (u);
\end{tikzpicture}
\hspace{\fill}
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (seq) at (0,0) {$PAR$};
   \node[provide] (p) at (-1,0) {};
   \node[use] (u) at (1,0) {$m$};

  \path[-]
    (p) edge node {} (seq)
    (seq) edge node {} (u);
\end{tikzpicture}
\hspace{\fill}
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (sync) at (0,0) {$SYNC$};
   \node[provide] (p) at (-1,0) {};
   \node[use] (u) at (1,0) {};

  \path[-]
    (p) edge node {} (sync)
    (sync) edge node {} (u);
\end{tikzpicture}
\\
a) Comp. SEQ
\hspace{\fill}
b) Comp. PAR
\hspace{\fill}
c) Comp. SYNC
\\
\caption{The three control components used by the back-end.}
\label{fig:ctrlcomponents}
\end{figure}

\begin{description}
\item[Sequence component (SEQ)] It is the direct representation of a sequence node of $\Gamma_{tsp}$. This component sequentially calls an ordered list of other components. It exposes an ordered \emph{use-multiple} port to be connected to the components to call in sequence.

\item[Parallel component (PAR)] It is the direct representation of a parallel node of $\Gamma_{tsp}$. This component simultaneously calls a set of other components. It exposes a \emph{use-multiple} port to be connected to the components to call in parallel.

\item[Synchronization component (SYNC)]. It is the direct representation to an update leaf of $\Gamma_{tsp}$. This component calls the synchronization of a given data. It exposes a \emph{use} port to be connected to the data to update.

\end{description}

%----------------------------------------
\subsection{Dump To Component Assembly}
%----------------------------------------
From $\Gamma_{tsp}$ and using the control components described above,
a direct dump can be done to a component assembly for an hybrid (data
and task) parallelization of a simulation. 
\fix{HC: ce qui suit est a changer avec la fusion du code de plusieurs composants}
However, it is also
possible to dump to a data parallelization only. In such a case the
computation of $\Gamma_{sync}$ is enough to generate the component
assembly. Actually a single SEQ component is thus needed and this
component is linked to all computations and synchronizations of
$\Gamma_{sync}$ directly.

Figure~\ref{fig:tsp:assembly} displays the assembly part corresponding to
$\Gamma_{tsp}$ of Figure~\ref{fig:tsp}. In this figure, the ports
linked to data (use and use-multiple ports of SYNC and K) are
represented but are not connected. Moreover, each computation
component is an instance of a specific K component using
the identification name of the computation in the MSL file.

\begin{figure*}[t]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   %seq0
   \node[component] (seq0) at (0,0) {$SEQ$};
   \node[provide] (seq0p) [left = 0.8cm of seq0] {};
   \node[use] (seq0u) [right = 1cm of seq0] {$m$};
   %c0
   \node[component] (c0) [right = 2cm of seq0u] {$K(c_0)$};
   \node[use] (c0u) [right = 0.8cm of c0] {};
   \node[right=0.2 of c0u] (star) {*};
   %sync0
   \node[component] (sync0) [below = 1cm of c0] {$SYNC$};
   \node[use] (sync0u) [right = 1cm of sync0] {$m$};
   %c1
   \node[component] (c1) [below = 0.8cm of sync0] {$K(c_1^*)$};
   \node[use] (c1u) [right = 0.8cm of c1] {};
   \node[right=0.2 of c1u] (star) {*};
   %par0
   \node[component] (par0) [below = 0.8cm of c1] {$PAR$};
   \node[use] (par0u) [right = 1cm of par0] {$m$};
  %seq1
   \node[component] (seq1) [right = 1cm of par0u] {$SEQ$};
   \node[use] (seq1u) [right = 1cm of seq1] {$m$};
   %par1
   \node[component] (par1) [right = 1cm of seq1u] {$PAR$};
   \node[use] (par1u) [right = 1cm of par1] {$m$};
  %c2
   \node[component] (c2) [right = 1cm of par1u] {$K(c_2)$};
   \node[use] (c2u) [right = 0.8cm of c2] {};
   \node[right=0.2 of c2u] (star) {*};
  %sync1
   \node[component] (sync1) [below = 0.8cm of c2] {$SYNC$};
   \node[use] (sync1u) [right = 1cm of sync1] {$m$};
  %c4
   \node[component] (c4) [below = 0.8cm of par1] {$K(c_4^*)$};
   \node[use] (c4u) [right = 0.8cm of c4] {};
   \node[right=0.2 of c4u] (star) {*};
  %c6
   \node[component] (c6) [below = 0.8cm of c4] {$K(c_6)$};
   \node[use] (c6u) [right = 0.8cm of c6] {};
   \node[right=0.2 of c6u] (star) {*};
  %seq2
   \node[component] (seq2) [below = 2.5cm of seq1] {$SEQ$};
   \node[use] (seq2u) [right = 1cm of seq2] {$m$};
  %c3
   \node[component] (c3) [right = 1cm of seq2u] {$K(c_3)$};
   \node[use] (c3u) [right = 0.8cm of c3] {};
   \node[right=0.2 of c3u] (star) {*};
  %c5
   \node[component] (c5) [below = 0.8cm of c3] {$K(c_5)$};
   \node[use] (c5u) [right = 0.8cm of c5] {};
   \node[right=0.2 of c5u] (star) {*};
  %c7
   \node[component] (c7) [below = 0.8cm of par0] {$K(c_7)$};
   \node[use] (c7u) [right = 0.8cm of c7] {};
   \node[right=0.2 of c7u] (star) {*};
  %sync2
   \node[component] (sync2) [below = 0.8cm of c7] {$SYNC$};
   \node[use] (sync2u) [right = 1cm of sync2] {$m$};
  %c8
   \node[component] (c8) [below = 0.8cm of sync2] {$K(c_8^*)$};
   \node[use] (c8u) [right = 0.8cm of c8] {};
   \node[right=0.2 of c8u] (star) {*};

   \path[-]
   %seq0
    (seq0) edge node {} (seq0u)
    (seq0p) edge node {} (seq0)
   %c0
    (seq0u) edge node {} (c0)
    (c0) edge node {} (c0u)
    %sync0
    (seq0u) edge node {} (sync0.west)
    (sync0) edge node {} (sync0u)
    %c1
    (seq0u) edge node {} (c1)
    (c1) edge node {} (c1u)
   %par0
    (seq0u) edge node {} (par0.west)
    (par0) edge node {} (par0u)
  %seq1
    (par0u) edge node {} (seq1)
    (seq1) edge node {} (seq1u)
  %par1
    (seq1u) edge node {} (par1.west)
    (par1) edge node {} (par1u)
  %c2
    (par1u) edge node {} (c2)
    (c2) edge node {} (c2u)
  %sync0
    (par1u) edge node {} (sync1.west)
    (sync1) edge node {} (sync1u)
  %c4
    (seq1u) edge node {} (c4.west)
    (c4) edge node {} (c4u)
  %c6
    (seq1u) edge node {} (c6.west)
    (c6) edge node {} (c6u)
  %seq2
    (par0u) edge node {} (seq2.west)
    (seq2) edge node {} (seq2u)
  %c3
    (seq2u) edge node {} (c3)
    (c3) edge node {} (c3u)
  %c6
    (seq2u) edge node {} (c5.west)
    (c5) edge node {} (c5u)
  %c6
    (seq0u) edge node {} (c7.west)
    (c7) edge node {} (c7u)
  %sync2
    (seq0u) edge node {} (sync2.west)
    (sync2) edge node {} (sync2u)
  %c8
    (seq0u) edge [bend right] node {} (c8.west)
    (c8) edge node {} (c8u)
   ;
\end{tikzpicture}
\caption{Component assembly representing the computation part generated from $\Gamma_{tsp}$ of Figure~\ref{fig:tsp}.}
\label{fig:tsp:assembly}
\end{center}
\end{figure*}

%----------------------------------------