### 2015-08-04 WOLFHPC15/article/ms_msc.tex
wip Julien

%----------------------------------------
\subsection{Multi Stencil Language}
%----------------------------------------

\begin{figure}[t]
\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++,numbers=left]
program ::= "mesh:" meshid 
            "mesh entities:" listmeshent
            "computation domains:" 
                       listcompdom
            "independent:"
                       listinde
            "data:" listdata
            "time:" iteration
            "computations:" listcomp
listmeshent ::= meshent listmeshent
             |  meshent
listcompdom ::= compdom listcompdom
             |  compdom
compdom ::= compdomid "in" listmeshent
listinde ::= inde listinde
          |  inde
inde ::= compdomid "and" compdomid
listdata ::= data listdata
          |  data
data ::= dataid "," meshent
iteration ::= num
listcomp ::= comp listcomp
          |  comp
comp ::= dataid "[" compdomid "]=" compid "(" 
            listdataread ")"
listdataread ::= dataread listdataread
              |  dataread
dataread ::= dataid "[" neighborid "]"
          |  dataid
\end{lstlisting}
\caption{Grammar of the MS language}
\label{fig:grammar}
\end{figure}

As mentioned in the previous section, the Multi Stencil Language (MSL) is an agnostic descriptive language for multi-stencil simulations. Six main sections are required in a MSL description that match the six elements of the formal definition; they describe: \textbf{1.}\,the mesh, \textbf{2.}\,the mesh entities, \textbf{3.}\,the computation domains with their inter-dependencies, \textbf{4.}\,the data, \textbf{5.}\,the time iterations, and finally \textbf{6.} the computations.
Figure~\ref{fig:grammar} shows the grammar of MSL. Lines 1 to 9 define what is a MSL program with the different parts mentioned above. The remaining of this section describes those different parts of a MSL program and presents an example.
We can note that, as MSL is a descriptive language, almost all terminals \texttt{meshid}, \texttt{meshdom}, \texttt{compdomid}, \texttt{dataid}, \texttt{compid} and \texttt{neighborid}, are string identifiers. Only \texttt{num} is an integer terminal.

\subsubsection*{Mesh and Mesh Entities}
As illustrated in the first line of the grammar in Figure~\ref{fig:grammar}, a mesh is simply defined by a string identifier.
Kind of mesh entities are described as a list of identifiers (lines 2, 10 and 11 in Figure~\ref{fig:grammar}).
Figure~\ref{fig:mslex} presents an example of a mesh named \texttt{cart} used to represent a Cartesian mesh, and three kind of mesh entities: \texttt{cell}, \texttt{edgex}, and \texttt{edgey}.

\subsubsection*{Computation Domains and Their Dependencies}
A computation domain is a subset of a given kind of entities from a mesh used during one or more computations.
For each computation domain, the kind of mesh entities on which it is mapped is specified (lines 3-4 and 12-14 on Figure~\ref{fig:grammar}).
By default two computation domains of the same entities are handled as intersecting except when an explicit independence relation is mentioned in the \emph{independent} section of the language (lines 5-6 and 15-17 in Figure~\ref{fig:grammar}).
On lines 4 to 6 of Figure~\ref{fig:mslex}, computation domains are defined for each kind of mesh entities: \texttt{allcell}, \texttt{alledgex} and \texttt{alledgey}; they are used to represent the whole domain.
On lines 7 and 8 two other domains of \texttt{edgex} are defined: \texttt{part1edgex} and \texttt{part2edgex}; they are used to represent the boundaries of the domain.
On line 10 the independence relation between \texttt{part1edgex} and \texttt{part2edgex} is specified.

\subsubsection*{Data and Time}
A data element is a quantity to simulate. It is mapped on a given kind of mesh entities, as illustrated in lines 7 and 18-20 of Figure~\ref{fig:grammar}. The \texttt{time} section simply indicates a number of iterations to perform in the simulation. In the example of Figure~\ref{fig:mslex} ten data elements are defined and 500 iterations (lines 11-22). %An extension could be to support a convergence criterion.

\subsubsection*{Computation Description}
The last part of the language contains the specifications of stencil kernels and local computations as defined in Section~\ref{sect:multistencil}.
This description does not contain a direct expression of the numerical expressions.
Instead, the term $exp$ of Equations~\ref{eq:st} and~\ref{eq:loc} simply corresponds to an identifier that references an implementation in an external language  (\cf Section~\ref{sect:component}).
The specification starts by the identifier of the computed data element ($w$ in the formal description) with the computation domain between brackets ($d$).
Then, after the equal sign, the kernel identifier ($exp$) is specified.
Finally, between the parenthesis is the list of identifiers of the data elements read by the computation with their stencil shape between brackets ($R$).
If the computation is local, no brackets appear ($R_l$). In the example of Figure~\ref{fig:mslex}, nine computations are described. On line 24, the data element \texttt{b} is computed on the domain \texttt{allcell} by the kernel \texttt{c\_0} which read data \texttt{a} without neighborhood shape.

\begin{figure}[t]
\begin{lstlisting}[basicstyle=\small,mathescape,frame=single,language=C++,numbers=left]
mesh: cart
mesh domains: cell,edgex,edgey
computation domains:
  allcell in cell
  alledgex in edgex
  alledgey in edgey
  part1edgex in edgex
  part2edgex in edgex
independent:
  part1edgex and part2edgex
data:
  a,cell
  b,cell
  c,edgex
  d,edgex
  e,edgey
  f,cell
  g,edgey
  h,edgex
  i,cell
  j,edgex
time:500
computations:
  b[allcell]=c0(a)
  c[alledgex]=c1(b[n1])
  d[alledgex]=c2(c)
  e[alledgey]=c3(c)
  f[allcell]=c4(d[n1])
  g[alledgey]=c5(e)
  h[alledgex]=c6(f)
  i[allcell]=c7(g,h)
  j[partedgex]=c8(i[n1])
\end{lstlisting}
\caption{Example of a MSL program}
\label{fig:mslex}
\end{figure}

%----------------------------------------
\subsection{MSC Parallelization}
%----------------------------------------
The MSC parallelization is a subpart of the overall compiler MSCAC.
It makes use of the ordered list of computations $\Gamma$, which can directly be extracted from the parser, to build a parallel representation of the computations of the overall multi-stencil program.
This parallelization phase is divided in five different steps to transform $\Gamma$ to a series-parallel tree decomposition~\cite{Valdes:1979:RSP:800135.804393}.
Data parallelism is handled in the first two steps, and task parallelism in the following three steps.
This section gives an overview of these five steps , but they are more formally defined with their associated algorithms in a research report~\cite{????}.

\subsubsection*{The Ordered List of Computations: $\Gamma$}
$\Gamma$ is directly obtained from the parser. Actually, the list of computations in the MSL program is already ordered: $\Gamma$ is a direct map of this list. In the example of Figure~\ref{fig:mslex}, $\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$.

\subsubsection*{The Synchronized Ordered List of Computations: $\Gamma_{sync}$}
In a data parallelization, data is split among processors and the same program is applied on each sub-part of the data.
This does however require additional synchronizations between processors.
For example, as a stencil kernel accesses neighbor values, values computed by another processor are needed and must be synchronized. 
The required synchronizations can be automatically computed from the ordered list of computations $\Gamma$.
A synchronization is needed each time a data read by a stencil computation has been written by a previous computation. In such a case, a computation is added before the stencil computation. This \emph{synchronization computation} reads the data to synchronize, and write the same data. The computation domain of such a synchronization computation encompasses the whole mesh entities on which the data is declared. As a result $\Gamma$ is transformed to a synchronized ordered list of computations $\Gamma_{sync}$.
For example, in Figure~\ref{fig:mslex}, the stencil computation $c_1$ reads the data element $b$ which has been written by $c_0$. For this reason the sublist $[c_0,c_1]$ of $\Gamma$ is transformed into $[c_0,sync_1,c_1]$ in $\Gamma_{sync}$. The new computation $sync_1$ reads and writes $b$ and is applied on the computation domain of $c_1$. As a result, a dependency is kept between $c_0$ and $sync_1$ and between $sync_1$, and $c_1$. The same transformation is performed for all stencils: $c_4$ and $c_8$. Thus, $\Gamma_{sync} = [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$.
\todo[inline]{JB: "is applied on the computation domain of $c_1$" vraiment, c'est pas ce que tu dis juste avant ...}

\subsubsection*{The Dependency Graph: $\Gamma_{dep}$}
From the synchronized ordered list of computations $\Gamma_{sync}$, a dependency graph $\Gamma_{dep}$ is built.
A dependency exists from computation $a$ to $b$ (including synchronizations) if and only if a data element read by $b$ has been written by $a$ in $\Gamma_{sync}$ with intersecting domains.
%In other words, the only case where the same data is manipulated by distinct kernels without inducing a dependency is when the computation domains do not intersect.
Nodes of the dependency graph represent computations of $\Gamma_{sync}$, while edges are dependencies between them. The dependency graph $\Gamma_{dep}$ is a directed acyclic graph (\emph{DAG}). For example, the dependency DAG $\Gamma_{dep}$ of the example of Figure~\ref{fig:mslex} is presented in Figure~\ref{fig:hyb}.

\begin{figure}[t]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node (c0) at (0,0) {$c_0$};
   \node[right=1 of c0] (sy1) {$sync_1$};
   \node[right=1 of sy1] (c1) {$c_1$};
   \node[above right=1 of c1] (c2) {$c_2$};
   \node[below right=1 of c1] (c3) {$c_3$};
   \node[right=1 of c2] (sy4) {$sync_4$};
   \node[right=1 of c3] (c5) {$c_5$};
   \node[right=1 of sy4] (c4) {$c_4$};
   \node[right=1 of c4] (c6) {$c_6$};
   \node[below right=1 of c6] (c7) {$c_7$};
   \node[right=1 of c7] (sy8) {$sync_8$};
   \node[right=1 of sy8] (c8) {$c_8$};

  \path[->]
    (c0) edge node {} (sy1)
    (sy1) edge node {} (c1)
    (c1)  edge node {} (c2)
          edge node {} (c3)
    (c2) edge node {} (sy4)
    (sy4) edge node {} (c4)
    (c4) edge node {} (c6)
    (c3) edge node {} (c5)
    (c5) edge node {} (c7)
    (c6) edge node {} (c7)
    (c7) edge node {} (sy8)
    (sy8) edge node {} (c8);
\end{tikzpicture}
\caption{$\Gamma_{hybrid}$ of the example of Figure~\ref{fig:mslex}}
\label{fig:hyb}
\end{center}
\end{figure}

\subsubsection*{The Minimal Series-Parallel Graph: $\Gamma_{msp}$}
Once a dependency graph is built, many solutions can be used to build a parallel application, as for example dynamic schedulers~\cite{Augonnet2011,Gautier:2013:XRS:2510661.2511383}. In this work a static scheduling of the dependency graph is built. To do so the dependency graph is transformed to a minimal series-parallel graph~\cite{Valdes:1979:RSP:800135.804393}. As it has been shown in~\cite{Valdes:1979:RSP:800135.804393}, the transitive reduction of a DAG is a minimal series-parallel graph if and only if a forbidden shape, called \emph{N-shape} illustrated in Figure~\ref{fig:n}, is not found in the DAG. Thus, to transform $\Gamma_{dep}$ to the minimal series-parallel graph $\Gamma_{msp}$, an algorithm~\cite{Mitchell:2004:CMV:1082101.1082117} is applied to remove all \emph{N-shapes} by an over-constraint, as illustrated in Figure~\ref{fig:over}.

\todo[inline]{JB: dans le DAG il y a une dependance de c2 a sync4, le graphe serie parallele les execute en parallele ...}

\begin{figure}[h!]
\begin{center}
\subfloat[][The forbidden $N$ shape.\label{fig:n}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (a) at (0,0) {$c_0$};
   \node[] (b) at (2,0) {$c_1$};
   \node[] (c) at (0,-1) {$c_2$};
   \node[] (d) at (2,-1) {$c_3$};

  \path[->]
    (a) edge node {} (b)
    (c) edge node {} (b)
        edge node {} (d);
  \end{tikzpicture}
  }
  \hspace{25pt}
  \subfloat[][Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$c_0$};
   \node[] (bc0) at (-1,0) {};
   \node[] (c1) at (2,0) {$c_1$};
   \node[] (c2) at (0,-1) {$c_2$};
   \node[] (bc2) at (-1,-1) {};
   \node[] (c3) at (2,-1) {$c_3$};

  \path[->]
    (bc0) edge [dotted] node {} (c0)
    (bc2) edge [dotted] node {} (c2)
    (c0) edge node {} (c1)
          edge [dashed] node [swap] {} (c3)
    (c2) edge node {} (c1)
        edge node {} (c3);
  \end{tikzpicture}
}
  \caption{Forbidden $N$ subgraph shape for a DAG to be minimal series-parallel.}
  \label{fig:forbidden}
\end{center}
\end{figure}

\subsubsection*{The Tree Decomposition: $\Gamma_{tsp}$}
Many works~\cite{Valdes:1979:RSP:800135.804393,Schoenmakers95anew} explains how to build a series-parallel tree decomposition from a minimal series-parallel graph. A series-parallel tree decomposition consists in the decomposition of the minimal series-parallel graph as a tree representing a set of \emph{sequences}, indicated by a $S$ node, and \emph{parallel} sections, indicated as a $P$ node. For example, the series-parallel tree decomposition $\Gamma_{tsp}$ of Figure~\ref{fig:hyb} is illustrated in Figure~\ref{fig:tsp}.

\begin{figure}[t]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0) at (0,0) {$\mathcal{S}$};
   \node[] (c0) at (-3,1) {$c_0$};
   \node[] (star1) at (-2,1) {$sync_1$};
   \node[] (c1) at (-1,1) {$c_1$};

   \node[] (p0) at (0,1) {$\mathcal{P}$};
   \node[] (s1) at (-1,2) {$\mathcal{S}$};
   \node[] (p1) at (-2,3) {$\mathcal{P}$};
   \node[] (c4) at (-1,3) {$c_4$};
   \node[] (c6) at (-0,3) {$c_6$};
   \node[] (c2) at (-2.5,4) {$c_2$};
   \node[] (star4) at (-1.5,4) {$sync_4$};
   \node[] (s2) at (1,2) {$\mathcal{S}$};
   \node[] (c3) at (0.5,3) {$c_3$};
   \node[] (c5) at (1.5,3) {$c_5$};

   \node[] (c7) at (1,1) {$c_7$};
   \node[] (star8) at (2,1) {$sync_8$};
   \node[] (c8) at (3,1) {$c_8$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (p1)
         edge node {} (c4)
         edge node {} (c6)
    (p1) edge node {} (c2)
         edge node {} (star4)
    (s2) edge node {} (c3)
         edge node {} (c5);
  \end{tikzpicture}
\caption{$\Gamma_{tsp}$ of Figure~\ref{fig:hyb}. Nodes $S$ represent sequences, nodes $P$ represent parallel sections.}
\label{fig:tsp}
\end{center}
\end{figure}

% \paragraph{Dump to a component assembly} Finally, from the series-parallel tree decomposition $\Gamma_{tsp}$ we could build a program using any parallel language containing \emph{parallel sections} and \emph{sequences of instructions} such as OpenMP~\cite{660313}, HPF~\cite{219857}, UPC~\cite{El-Ghazawi:2006:UUP:1188455.1188483} etc. In this paper we study the dump of $\Gamma_{tsp}$ to a component assembly which needs the introduction of \emph{control components}, described in the next section.