### 2015-08-06 WOLFHPC15/article/conclusion.tex

Conclusion :
\begin{itemize}
\item Multi-stencil formalism
\item MSL and MSC for data and hybrid automatic parallelization
\item Discussion dump to components
\item Evaluation
\end{itemize}

Perspectives :
Improve MSL for multiple meshes = multi-grids solvers + hybrid meshes\\
Evaluate component-based runtime :
\begin{itemize}
\item Global Arrays instead of SKelGIS
\item Dump to OpenMP for comparison
\item MSC to dynamic scheduler (starPU, XKaapi)
\end{itemize}