### 2015-08-21 WOLFHPC15/article/article.tex
fix abstract

\documentclass{acm_proc_article-sp}
\usepackage{tikz}
\usepackage{listings}
\usepackage{url}
\usepackage{cite}
\usepackage{subfig}
\usepackage{filecontents}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{todonotes}

%----------TIKZ
\usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}

\tikzset{
    use/.style={
    circle,draw=black,fill=black,scale=0.5,text=white
    },
    mpi/.style={
    rectangle,draw=black,fill=black,scale=0.8,text=white
    },
    provide/.style={
    circle,draw=black,fill=white,scale=0.5
    },
    component/.style={
    rectangle,rounded corners=3pt,draw=black
    },
    dcomponent/.style={
    rectangle,rounded corners=3pt,dashed,draw=black
    },
}
%-------------
\newenvironment{mydef}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\fix}[1]{\todo[inline,color=red!80]{#1}}
% \renewcommand{\todo}[2][]{}

\newcommand{\cf}{\emph{cf.}\xspace}
\newcommand{\ie}{\emph{i.e.,}\xspace}
\newcommand{\llc}{L\textsuperscript{2}C\xspace}

\begin{document}

\title{From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study}

\numberofauthors{3} 
\author{
% 1st. author
\alignauthor
Julien Bigot\\
       \affaddr{CEA, Maison de la Simulation}\\
       \affaddr{USR 3441, CEA Saclay}\\
       \affaddr{91191 Gif-sur-Yvette, France}\\ % remove cedex
       \email{julien.bigot@cea.fr}
% 1st. author
\alignauthor
H\'el\`ene Coullon\\
       \affaddr{INRIA, LIP}\\
       \affaddr{ENS Lyon, 46 All\'ee d'Italie}\\
       \affaddr{Lyon, France}\\
       \email{helene.coullon@inria.fr}
% 2nd. author
\and
\alignauthor
Christian P\'erez\\
       \affaddr{INRIA, LIP}\\
       \affaddr{ENS Lyon, 46 All\'ee d'Italie}\\
       \affaddr{Lyon, France}\\
       \email{christian.perez@inria.fr}
}

\maketitle

%----------------------------------------
\begin{abstract}
  High performance architectures evolve continuously to be more powerful. Such architectures also usually become more difficult to use efficiently. As a scientist is not a low level and high performance programming expert, Domain Specific Languages (DSLs) are a promising solution to automatically and efficiently write high performance codes. However, if DSLs ease programming for scientists, maintainability and portability issues are deported from scientists to DSL designers. This paper deals with an approach to improve maintainability and programming productivity of DSLs through the generation of a component-based parallel runtime. To study it, the paper presents a  DSL for \emph{multi-stencil} programs, that is evaluated on a real-case of shallow water equations.
\end{abstract}
%----------------------------------------

%\keywords{Component models, domain specific languages, multi-stencil, numerical simulations} % NOT required for Proceedings

%----------------------------------------
\section{Introduction}
\label{sect:intro}
\input{introduction.tex}
%----------------------------------------
\section{Stencil and Multi-Stencil}
\label{sect:concept}
\input{concept.tex}
%----------------------------------------
\section{MSL and MSCAC Overview }
\label{sect:mscac}
\input{mscac.tex}
%----------------------------------------
\section{Language and Parallelization Description}
\label{sect:msmsc}
\input{ms_msc.tex}
%----------------------------------------
\section{A Component based Back-End}
\label{sect:component}
\input{component.tex}
%----------------------------------------
\section{Real-case Evaluation}
\label{sect:eval}
\input{evaluation.tex}
%----------------------------------------
\section{Related Work}
\label{sect:related}
\input{related.tex}
%----------------------------------------
\section{Conclusion}
\label{sect:conclusion}
\input{conclusion.tex}
%----------------------------------------

\section{Acknowledgments}

We acknowledge Grid'5000 for evaluations.

\small
\bibliographystyle{abbrv}
\bibliography{biblio}

\end{document}