### 2016-02-01 ARTICLES/IJPP/article/msl.tex

\begin{filecontents*}{grammar.txt}
program ::= "mesh:" meshid 
            "mesh entities:" listmeshent
            "computation domains:" 
                      listcompdom
            "independent:"
                      listinde
            "stencil shapes:"
                      liststencil
            "data:" listdata
            "scalar" listscalar
            listloop

listmeshent ::= meshent listmeshent 
                        |  meshent
listcompdom ::= compdom listcompdom 
                        |  compdom
compdom ::= compdomid "in" listmeshent
listinde ::= inde listinde |  inde
inde ::= compdomid "and" compdomid
liststencil ::= stencil liststencil 
                        | stencil
stencil ::= stencilid "from" meshent "to" meshent
listdata ::= data listdata |  data
data ::= dataid "," meshent
listscalar ::= scalar listscalar 
                      | scalar

listloop ::= loop listloop | loop
loop ::=  "time:" iteration
          "computations:" listcomp

iteration ::= num | criteria
criteria ::= criteriaid "(" listscalarread ")"
listscalarread ::= scalar listscalarread 
                          |  scalar
listcomp ::= comp listcomp |  comp
comp ::= dataid "[" compdomid "]=" 
                    compid "({" 
                    listscalarread "},{" 
                    listdataread "})"
listdataread ::= dataread listdataread 
                          |  dataread
dataread ::= dataid "[" stencil "]" 
              |  dataid
\end{filecontents*}

\begin{figure}[!h]
  \hspace{5mm}
  \begin{minipage}[!h]{0.5\textwidth}
    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left]{grammar.txt}}   
    \caption{Grammar of the Multi-Stencil Language. \label{fig:grammar}}
  \end{minipage}
\end{figure}

\begin{filecontents*}{exmsl.txt}
mesh : cart
mesh entities : cell, edgex
computation domains :
  d1 in cell
  d2 in edgex
independent :
  d1 and d2
stencil shapes : 
  ncc from cell to cell
  nce from cell to edgex
  nec from edgex to cell
data :
  A,cell
  B,cell
  C,edgex
  D,cell
  E,cell
  F,cell
  G,cell
  H,edgex
  I,cell
  J,cell
scalar : mu, tau
time : 500
computations :
  B[d1] = k0({tau},{A})
  C[d2] = k1({},{B[n2]})
  D[d1] = k2({},{C})
  E[d1] = k3({},{C})
  F[d1] = k4({},{D,C[n3]})
  G[d1] = k0({mu,tau},{E})
  H[d2] = k6({},{F})
  I[d1] = k7({},{G,H})
  J[d1] = k8({mu},{I[n1]})
\end{filecontents*}

\begin{figure}[!h]
  \hspace{5mm}
  \begin{minipage}[!h]{0.4\textwidth}
    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=right]{exmsl.txt}}   
    \caption{Example of program using the Multi-Stencil Language. \label{fig:exmsl}}
  \end{minipage}
\end{figure}