### 2016-05-27 ARTICLES/IJPP/article/introduction.tex
IJPP article introduction

Modern high performance architectures are becoming more and more heterogeneous and difficult to use. The current world's top supercomputer Tianhe-2~\footnote{url{www.top500.org}} is composed of multi-cores processors and accelerators, and is able to reach a peak performance of thirty floating point operations par seconds. Heterogeneity and complexity of architectures will probably continue to increase in future.

One the big challenges to solve to be able to use exascale computers is to propose programming models which gives access to high performance computing (HPC) to many scientists and not only to a few HPC specialists~\cite{ETP4HPC2013}. Actually, applications which runs on supercomputers and which needs such a computation power are physics, weather or genomic applications, which are not implemented by HPC specialists most of the time.

Many low level execution models for exascale computers focuss onto the dynamic scheduling of task graphs combined to message passing models to be able to use more than one machine~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011}. Those models increase HPC code portability and reach an interesting performance onto heterogeneous architectures. At a higher abstraction level, general purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} follow the direction of task graph scheduling proposed by those execution models. However, for non-experts end-users, general purpose languages still are difficult to use and to tune for a given application onto a given architecture. The current easiest (with the higher abstraction level), but still efficient, programming model is Domain Specific Language (DSL). Such a language is specific to the end-user domain and proposes a grammar which is easy to understand and completely sequential. The DSL compiler is able, by knowing the specific domain targetted, to automatically apply parallelization and specific optimizations to produce a high performance back-end code. Thus, a DSL is able to split end-user concerns from HPC concerns which is relevant to reach exascale programming models.

However, many DSLs have been proposed for many domains yet, but none of them has been able to reuse work already done by another language. In other words, software engineering properties have to be integrated into DSL conception, such that a new DSL can be seen as a composition of parallelization, optimizations or even languages semantics already proposed by others.

For example, to numerically solve a set of partial differential equations (PDEs), iterative methods are frequently used to approximate the exact solution through a discretized phenomena. A very well known and usual way to discretize PDEs is to transform them to explicit numerical schemes, also often called \emph{stencils}. Many DSLs have been proposed for stencil computations~\cite{spaaTangCKLL11,citeulike12258902,Ragan-Kelley:2013:HLC:2491956.2462176,DeVito:2011:LDS:2063384.2063396,Camier:2015:IPP:2820083.2820107}, as it will be described in Section~\ref{sect:rel}. Many of them use same kind of parallelization, data structures or optimizations, however each one has been built from scratch to deal with another additionnal specific case.

We present the Multi-Stencil Language (MSL) DSL for stencil-based numerical simulations. MSL is a language with a light grammar to only describe such a numerical simulation without implementation details. From the description, the compiler has enough information to extract and build an empty parallel pattern of the simulation, which can be filled, in a second step, by implementation concerns. Te parallel pattern generated by the language is able to use different existing languages and libraries as it is independent from implementation choices. Moreover, the parallelization performed by the language is large enough to be compatible with many architectures and back-end languages. Contributions presented in this paper are : the computational model of a multi-stencil program and its parallelization formalism; the MSL grammar and its compiler; and a performance evaluation of a real case numerical simulation up to 16.384 cores.

Section~\ref{sect:rel} introduces related work on stencils DSLs. Section~\ref{sect:formalism} formally explains the targetted domain and its computational model. From this model can be extracted the grammar of MSL in Section~\ref{sect:msl}. Section~\ref{sect:parallelism} shows how parallelism can be extracted from this light grammar of MSL. Sections~\ref{sect:msp} and~\ref{sect:comp} detail choices that have been done in this paper to evaluate MSL, and Section~\ref{sect:eval} shows performance results of the language. Finally, Section~\ref{sect:concl} concludes and proposes perspectives on this work.