### 2016-05-30 ARTICLES/IJPP/article/evaluation.tex
eval

This section evaluates the MSL compiler and the code it generates.
We first evaluate the generated code in the case where only data parallelization and fusion are applied and compare this to a full SkelGIS program.
Then, we analyze a case taking advantage of the full the hybrid parallelization combining data and task parallelism and validate the performance model proposed in Section~\ref{sect:tsp}.

All evaluations presented in this section are based on a real case study of the shallow-Water Equations as solved in  the FullSWOF2D\footnote{\url{http://www.univ-orleans.fr/mapmo/soft/FullSWOF/}}~\cite{Ferrari2004,CPE:CPE3494} code from the MAPMO laboratory, University of Orl\'eans.
We have developed a MSL version of FullSWOF2D that contains 3~mesh entities, 7~computation domains, 48~data and 98~computations (32~stencil kernels and 66~local kernels).

%-------------------------------------
\subsection{Compiler}

Table~\ref{fig:exectime} illustrates the execution time of each step of the MSL compiler for the FullSWOF2D example.
This has been computed on a laptop with a dual-core Intel Core i5 1.4~GHz, and 8~GB of DDR3.
While the overall time of 4.6 seconds remains reasonable, one can notice that the computation of the $TSP$ tree is by far the longest step.
As a matter of fact, the complexity of the algorithm for N-shapes removal is $O(n^3)$.
The replacement of the static scheduling by a direct scheduling of the DAG by dedicated tools should solve this in the future.

\begin{table}[!h]
 \begin{center}
 \begin{tabular}{|c|c|c|c|c|}
  \hline
   Step & Parser & $\Gamma_{sync}$ & $\Gamma_{dep}$ & $TSP$\\
   \hline
   Time (ms) & 1 & 2 & 4.2 & 3998.5\\
   \hline
   \% & 0.022 & 0.043 & 0.09 & 86.6\\
   \hline
 \end{tabular}
\caption{Execution times of the MSL compiler}
\label{fig:exectime}
 \end{center}
\end{table}

%-------------------------------------
\subsection{Data parallelism}

In this evaluation, we disable task-parallelism (fix its degree to 1) so as to focus on data-parallelism.
In the implementation evaluated in this paper, the data parallelism is handled using SkelGIS that has already been evaluated in comparison with a native MPI implementation for the FullSWOF2D example~\cite{CPE:CPE3494}.
In addition to SkelGIS, MSL automatically inserts the required MPI synchronizations and identifies potential fusion of kernels to reduce cache misses.

\begin{table}[!ht]
\begin{center}
 \begin{tabular}{|c|c|}
   \hline
    Cluster & \textbf{TGCC Curie Thin Nodes}\\
     \hline         
    Processor & 2$\times$SandyBridge\\
    & (2.7 GHz)\\
    Cores/node & 16 \\
    RAM/node & 64 GB\\
    RAM/core & 4GB\\
    Compiler [-O3] & gcc 4.9.1\\
    MPI & Bullxmpi\\
    Network & fat-tree Infiniband\\
    \hline
 \end{tabular}
 \caption{\label{tab:TGCC}Hardware configuration of TGCC Curie Thin nodes.}
 \end{center}
\end{table}

This section compares the performance of the code produced by MSL for FullSWOF2D with a plain SkelGIS program where synchronizations and fusion choices have been done by the developer without any automatic support.
This enables to evaluate both the choices made by the compiler as well as the potential overheads of using \llc~\cite{l2c} that is not used in the plain SkelGIS version.
The evaluations have been performed on the Curie supercomputer (TGCC, France) described in Table~\ref{tab:TGCC}.
Each evaluation has been performed ten times and the median is presented in results.

\paragraph{\textbf{Weak scaling}}
\begin{figure}[t]
\begin{minipage}{.475\textwidth}
  \includegraphics[width=\textwidth]{../results/weak_scaling/400/median_weak.pdf}
  \caption{weak-scaling with $400 \times 400$ domain per core.}
  \label{fig:weak1}
\end{minipage}
\hfill
\begin{minipage}{.475\textwidth}
  \includegraphics[width=\textwidth]{../results/weak_scaling/800/median_weak.pdf}
  \caption{weak-scaling with $800 \times 800$ domain per core.}
  \label{fig:weak2}
\end{minipage}\end{figure}
%In a weak scaling, a constant amount of work is given to each core while increasing the overall number of cores.
%As a result, the overall domain size also increase with the number of cores.
%A weak scaling is a good evaluation to detect overheads which are independent from the amount of computation performed by each core, such as collective communications, or overheads introduced by \llc components, for example.
Figures~\ref{fig:weak1} and~\ref{fig:weak2} show a weak scaling for a $400 \times 400$ points by core and a $800 \times 800$ points by core from 16 cores to 16.384 cores.
Minimum and maximum values are also shown as error bars in figures.

One can notice that the MSL code produces a slightly better execution time on the $400 \times 400$ domain and slightly worse a small overhead on the $800 \times 800$ domain.
This might be explained by slightly different compiler optimization flags since \llc requires compilation as dynamic libraries with the \emph{-fpic} compilation flag while the SkelGIS version does not.
This flag can have slight positive or negative effect on code performance depending on the situation and might be responsible for the observed difference.
In any case, these difference are barely noticeable and often included in the error bar of the measure and both version globally perform similarly.
The MPI synchronizations performed by MSL perform as good as manually specified ones at scale and \llc does not introduce any noticeable overhead.

\paragraph{\textbf{Strong scaling}}
\begin{figure}[t]\begin{center}
  \includegraphics[width=.6\textwidth]{../results/strong_scaling/10K_1K/median_strong.pdf}
  \caption{Strong scaling on a $10k \times 10k$ domain.}
  \label{fig:strong}
\end{center}\end{figure}
% The second evaluation is a strong scaling evaluation.
% Unlike a weak scaling, a strong scaling keeps a constant overall size of problem, while the number of cores increase.
% As a result, the sub-domain computed by each core reduces while the number of cores increases.
% The strong scaling is a good additional evaluation to the weak scaling, showing different kinds of overheads, such as cache misses (when the size of sub-domain becomes small enough to fit into cache memory), or small constant overheads.
Figure~\ref{fig:strong} shows the number of iteration per second for a 10k$\times$10k global domain size from 16 to 16.384 cores.
The ideal strong scaling is also illustrated.

First, one can notice that the strong scaling evaluated for the code generated by MSL is close to the ideal speedup up to 16.384 cores, which is a very good result. Moreover, no overheads are introduced by MSL which shows that automatic synchronization detections and automatic fusion detections are the same one that the one written manually into the SkelGIS code of FullSWOF2D. Finally, no overheads are introduced by components of \llc. A small behavior difference can be noticed with $2^9=512$ cores, however this variation is no longer observed with 1024 cores.

\paragraph{\textbf{Fusion}} Finally, to evaluate the data parallelization technique automatically introduced by MSL, the fusion optimization is evaluated.
Figure~\ref{fig:fusion} shows the number of iterations per second as a function of the number of cores, for FullSWOF2D with and without fusion optimization and onto a $500 \time 500$ domain. As explained in Section~\ref{sect:fusion}, the MSL fusion happens at a high level and is most of the time done naturally by a computer scientist. However, for a non computer scientist which write its numerical codes, an automatic proposition of such fusions makes the implementation easier. Moreover, one can notice that the performance is clearly improved (around 40\%) by this fusion.

\begin{figure}[!h]\begin{center}
  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/fusVSbase.pdf}
  \caption{Strong scaling on a 500x500 domain, with and without the fusion optimization.}
  \label{fig:fusion}
\end{center}\end{figure}

%-------------------------------------
\subsection{Hybrid parallelism}

\begin{table}[th]
 \begin{center}
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline 
   Level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
   \hline
   Frequency & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
   \hline
 \end{tabular}
\caption{Parallelism level (number of parallel tasks) and the number of times this level appears.}
\label{fig:freq}
 \end{center}
\end{table}

In this evaluation, we introduce task parallelism to evaluate the hybrid parallelization offered by MSL that mixes data parallelization and task parallelization.
In the implementation evaluated in this paper and in addition to SkelGIS, the hybrid parallelization relies on OpenMP.

The series-parallel tree decomposition $TSP$ of this simulation, extracted by the MSL compiler, is composed of 17~sequence nodes and 18~parallel nodes. Table~\ref{fig:freq} represents the number of time a given level of parallelism, \ie the number of tasks to perform concurrently, is observed in the final tree. One can notice that the level of task parallelism extracted from the Shallow water equations is limited by two sequential parts in the application (level 1). Moreover, a level of 16~parallel tasks is reached two times, and five times for the fourth level.
This means that if two threads are dedicated to task parallelism, two parts of the code at least will not take advantage of this and that no part of the code would benefit from more than 16 threads.
The task parallelism alone is therefore clearly not enough to take advantage of even a single node of a cluster that typically supports today between 16 and 32 threads (16 for Curie).

\begin{figure}[bh]\begin{center}
  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/analytic/times.pdf}
  \caption{Computation vs communication times in the data parallelization technique.}
  \label{fig:limit}
\end{center}\end{figure}

On the other hand, Figure~\ref{fig:limit} illustrates limitations of data parallelization technique alone.
While the computation time decreases linearly with the number of core used, the communication behavior is much more erratic.
Between 2 and 16 cores, the communication happen inside a single node and the time is small and nearly constant.
There is a small oscillation that might be explained by difference between number of processors that are power of 4 where the communications are all the same and those that are not where there is an imbalance.
Starting with 32 cores, there are multiple nodes and the communication time is typically modeled as $L+S/B$ where $L$ is the latency, $S$ the data size and $B$ the bandwidth.
This explains the decrease of time from 32 to 128 cores where the data sizes communicated by each process decreases.
The increased observed after that might be due to the fact that with the increased number of processes the fat-tree becomes deeper and the latencies increase.

\begin{figure}[th]\begin{center}
  \includegraphics[width=.6\textwidth]{../results/task_scaling/500_200/base_close_median.pdf}
  \caption{Strong scaling comparisons between data parallelization and hybrid parallelization. A \emph{close} OpenMP clause is used to bind threads onto cores.}
  \label{fig:close}
\end{center}\end{figure}

All in all, when the number of core used increases, the computation/communication ratio becomes poorer and poorer.
As a result, the data parallelism alone fails to provide enough parallelism to leverage the whole machine and other sources of parallelism have to be found.
The blue curve of Figure~\ref{fig:close} shows the total execution time for the same example as in Figure~\ref{fig:limit} and as expected, the speedup bends down from 256 to 2048 cores.

In addition to the blue curve, Figure~\ref{fig:close} shows strong scalings for the same example but using an hybrid parallelization. For example, the purple curve shows the parallelization which uses 8 cores, to perform the task parallelization, for each process used for data parallelization (\ie MPI process). As a result, for example, when using 2 machines of the TGCC cluster, with a total of 32 cores, 4 cores are used for SkelGIS MPI processes, for data parallelization, and for each one 8 cores are used for task parallelization ($4 \times 8 = 32$).

As a result, and as explained in Section~\ref{sect:perfs}, data is less divided into sub-domains and the effect which is observe onto the blue curve is delayed. This figure shows a comparison with 2, 4, 8 and 16 cores per MPI process for task parallelization.

From 2 to 8 cores, the improvement of the strong scaling is clear. However, reaching 16 cores, an important initial overhead appears and in addition to this, the curve bends down rapidly instead of improving the one with 8 cores for task parallelization. Two different phenomena happen.

First, thin nodes of the TGCC Curie are built with two NUMA nodes each of 8 cores. As a result, when increasing the number of OpenMP cores for task parallelization from 8 to 16 cores, an overhead is introduced by exchanges of data between memories of the two NUMA nodes. This phenomena is illutrated in Figure~\ref{fig:spread}. In this figure, a different strategy is used to bind threads onto available cores (using OpenMP). This strategy, called \emph{spread} instead of \emph{close} in Figure~\ref{fig:close}, binds threads on cores in order to spread as much as possible onto resources, which means that the two NUMA nodes are directly used whatever the number of cores kept for task parallelization. As a result, and as shown in the figure, using 2, 4 and 8 cores an initial overhead is introduced.

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_spread_median.pdf}}
  \caption{Strong scaling comparisons between data parallelization and hybrid parallelization. A \emph{spread} OpenMP clause is used to bind threads onto cores.}
  \label{fig:spread}
\end{center}\end{figure}

The second phenomena, which happens in Figure~\ref{fig:close} with 16 cores for tasks parallelization, is due to the level of parallelization introduced by the task parallelization technique. Actually, as illustrated in Table~\ref{fig:freq}, only two steps of the $TSP$ static scheduling generated by the MSL compiler can take advantage of 16 cores among a total of 18 steps. This phenomena has been explained in Section~\ref{sect:perfs} by the variable $F_{task}$ and the fact that it is not always true that $F_{task}=P_{task}$. This explains why using 16 cores for task parallelization in Figure~\ref{fig:spread} is still less efficient than using 8 cores even if the two NUMA nodes are always used in this evaluation.

Finally, to completely validate the performance model introduced in Section~\ref{sect:perfs}, and to understand when the hybrid parallelization becomes more interesting than the data parallelization, Figure~\ref{fig:tth2} represents $T_{COM1}$ and $T_{COM2}+T_{task}$ of Equation~(\ref{eq:hyb}), for the best case, \ie when 8 cores are used for task parallelization, and with a \emph{close} OpenMP bind of threads onto cores. Table~\ref{fig:tth} gives time details.

\begin{table}[!h]
 \begin{center}
 \begin{tabular}{|c|c|c|c|c|}
    \hline 
    & $T_{COM1}$ & $T_{COM2}$ & $T_{task}$ & Equation~(\ref{eq:hyb})\\
   \hline
   16 cores ($2 \times 8$) & 0.0005 & 0.00032 & 0.013 & False\\
   32 cores ($4 \times 8$) & 0.0018 & 0.00045 & 0.0062 & False\\
   64 cores ($8 \times 8$) & 0.0013 & 0.00038 & 0.0034 & False\\
   128 cores ($16 \times 8$) & 0.00075 & 0.0005 & 0.0023 & False\\
   256 cores ($32 \times 8$) & 0.00077 & 0.0018 & 0.001 & False\\
   512 cores ($64 \times 8$) & 0.0029 & 0.0013 & 0.00052 & True\\
   1024 cores ($128 \times 8$) & 0.018 & 0.00075 & 0.00029 & True\\
   2048 cores ($256 \times 8$) & 0.0623 & 0.00077 & 0.00016 & True\\
   \hline
 \end{tabular}
\caption{Execution times (seconds) of $T_{COM1}$, $T_{COM2}$ and $T_{task}$ for 8 cores for task parallelization. Verification of the Equation~(\ref{eq:hyb}).}
\label{fig:tth}
 \end{center}
\end{table}

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/tth.pdf}}
  \caption{Execution times (seconds) of $T_{COM1}$ and $T_{COM2} + T_{task}$ for 8 cores for task parallelization. Verification of the Equation~(\ref{eq:hyb}).}
  \label{fig:tth2}
\end{center}\end{figure}

Figure~\ref{fig:tth2} and Table~\ref{fig:tth} perfectly matches results observed in Figure~\ref{fig:close} for 8 cores used for task parallelization per core used for data parallelization. As a result, the hybrid parallelization becomes better with a total of 512 cores.