### 2016-11-08 ARTICLES/OTHER/C2S_EXA/slides.tex
final

\documentclass{beamer}

\usepackage{beamerthemeDresden}
\usepackage[utf8]{inputenc}
\usecolortheme{beaver}

\usepackage{listings}
\usepackage{xcolor}
\lstdefinestyle{base}{
  basicstyle=\ttfamily\color{black} \footnotesize,
  moredelim=**[is][\color{blue}]{@}{@},
  moredelim=**[is][\color{red}]{&}{&}
}

\setbeamertemplate{navigation symbols}{
\insertframenumber/
\inserttotalframenumber
}

\usepackage{color}
\definecolor{vert}{rgb}{0,0.5,0}
\definecolor{violet}{rgb}{0.5,0,0.5}

\usepackage{tikz}
\usetikzlibrary{calc,positioning,matrix,arrows,shapes.geometric,shapes.symbols,shapes.misc,shapes,automata,petri,decorations.markings,shadows}
\usepackage{array}
\usepackage{subfig}

%----------TIKZ
\usetikzlibrary{calc,arrows,shapes,automata,petri,positioning,decorations.markings,shadows}

\tikzset{
    use/.style={
    circle,draw=black,fill=black,scale=0.5,text=white
    },
    mpi/.style={
    rectangle,draw=black,fill=black,scale=0.8,text=white
    },
    provide/.style={
    circle,draw=black,fill=white,scale=0.5
    },
    component/.style={
    rectangle,rounded corners=3pt,draw=black
    },
    dcomponent/.style={
    rectangle,rounded corners=3pt,dashed,draw=black
    },
    progm/.style={rectangle,dashed, draw=black, thin, text width=10em, text centered,rounded corners, minimum height=2em},
  line/.style={draw, thin, ->, shorten >=2pt},
  purp/.style={rectangle, draw=violet,fill=violet!20, text width=10em, text centered,thin,rounded corners,inner sep=3pt},
  purp2/.style={rectangle, draw=violet,fill=violet!20, text width=15em, text centered,thin,rounded corners,inner sep=3pt},
  progbl/.style={rectangle, draw=blue, thin, text width=7em, text centered,rounded corners, minimum height=2em},
  progrg/.style={rectangle, draw=red, thin, text width=7em, text centered,rounded corners, minimum height=2em},
  proggr/.style={rectangle, draw=vert, thin, text width=7em, text centered,rounded corners, minimum height=2em},
  progbrg/.style={rectangle, draw=red, very thick, text width=10em, text centered,rounded corners, minimum height=2em}
}
%-------------

\def\pprec{\mathrel{\scalebox{.9}[1]{$\prec$}\mkern-3mu%
  \scalebox{.4}[1]{$\prec$}\mkern-5.5mu\scalebox{.4}[1]{$\prec$}}}

\makeatletter
    \newenvironment{withoutheadline}{
        \setbeamertemplate{headline}[default]
        \def\beamer@entrycode{\vspace*{-\headheight}}
    }{}
\makeatother

\AtBeginSection[]
{
\begin{frame}{Table of Contents}
\tableofcontents[currentsection]
\end{frame}
}
%-------------------------------------------------------------------
\title[The Multi-Stencil Language]{Using Component Model to implement a DSL: The Multi-Stencil Language Case Study}
\author[]{Hélène Coullon, Julien Bigot, \underline{Christian Perez}} % [Hélène Coullon (INRIA), Julien Bigot (CEA), Christian Perez (INRIA)]
\institute[]{Avalon project team\\Maison de la simulation}
\date{C2S@Exa -- Scientific Days \\Paris, November 8th, 2016}
%-------------------------------------------------------------------

\begin{document}
%gets rid of bottom navigation bars
\setbeamertemplate{footline}[page number]{}

%gets rid of navigation symbols
\setbeamertemplate{navigation symbols}{}
%---------------

\begin{frame}
    \titlepage
    \vspace{1.25cm}
    \resizebox{2.5cm}{!}{\includegraphics{images/logo-inria.png}}\hspace{1cm}
    \resizebox{1cm}{!}{\includegraphics{images/logo-mdls.png}}\hfill
    \resizebox{1cm}{!}{\includegraphics{images/elci_logo.png}}
\end{frame}

\begin{frame}{Table of Contents}
  \tableofcontents
\end{frame}

%% \begin{frame}
%%   \begin{enumerate}
%%   \item On // langage => low level vs high level => a (temporary) solution: DSL ?
%%   \item On DSL (2 slides)
%%   \item how to productivively build efficient DSL (MDS + component) and their runtime
%%   \item this talk: a case study
%%   \item MSL overview
%%   \item Compiler overview
%%   \item runtime overview
%%   \item eval de perf
%%   \item conclusion
%%   \end{enumerate}
%% \end{frame}

%-------------------------------------------------------------
% INTRODUCTION / MOTIVATION
%-------------------------------------------------------------
\section{Introduction}
\begin{frame}
\frametitle{High performance computing and parallelism in 2016}
\begin{center}
\only<1>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Programming models}; \\
      \node[progrg] (mod) {Libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
    \end{scope}
  \end{tikzpicture}
}
%--------
\only<2>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Programming models}; \\
      \node[progrg] (mod) {Libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
    \end{scope}
  \end{tikzpicture}
}
%--------
\only<3>{

  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Programming models}; \\
      \node[progrg] (mod) {Libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \node[fill=white,right of=mod,node distance=3cm] (bsp) {\scriptsize \textcolor{red}{BSP}};
    \node[fill=white,right of=bsp,node distance=1.5cm] (mpi) {\scriptsize \textcolor{red}{MPI}};
    \node[fill=white,right of=mpi,node distance=1.5cm] (openmp) {\scriptsize \textcolor{red}{OpenMP}};
    \node[fill=white,right of=openmp,node distance=1.5cm] (cuda) {\scriptsize \textcolor{red}{Cuda}};
    \node[fill=white,right of=cuda,node distance=1.5cm] (opencl) {\scriptsize \textcolor{red}{OpenCL ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
      \path (mpi) edge[->,thin, bend right=50] (openmp);
      \path (mpi) edge[->,thin, bend right=50] (cuda);
      \path (bsp) edge[->,thin, bend right=50] (mpi);
    \end{scope}

  \end{tikzpicture}
}
%--------
\only<4>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Programming models}; \\
      \node[progrg] (mod) {Libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \node[fill=white,right of=mod,node distance=3cm] (bsp) {\scriptsize \textcolor{red}{BSP}};
    \node[fill=white,right of=bsp,node distance=1.5cm] (mpi) {\scriptsize \textcolor{red}{MPI}};
    \node[fill=white,right of=mpi,node distance=1.5cm] (openmp) {\scriptsize \textcolor{red}{OpenMP}};
    \node[fill=white,right of=openmp,node distance=1.5cm] (cuda) {\scriptsize \textcolor{red}{Cuda}};
    \node[fill=white,right of=cuda,node distance=1.5cm] (opencl) {\scriptsize \textcolor{red}{OpenCL ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
      \path (mpi) edge[->,thin, bend right=50] (openmp);
      \path (mpi) edge[->,thin, bend right=50] (cuda);
      \path (bsp) edge[->,thin, bend right=50] (mpi);

      \path (mpi) -- (donn);
      \path (mpi) -- (tach);
      \path (mpi) -- (mess);
      \path (bsp) -- (donn);
      \path (bsp) -- (mess);
      \path (openmp) -- (tach);
      \path (openmp) -- (donn);
      \path (cuda) -- (tach);
      \path (opencl) -- (tach);
      \path (opencl) -- (mess);
    \end{scope}
  \end{tikzpicture}
}
%---------
\only<5>{
  \begin{tikzpicture}
    \matrix[row sep=0.6cm,column sep=0.5cm] {
      \node[progbl] (para) {Programming models}; \\
      \node[progrg] (mod) {Libraries\\ and languages};\\
      \node[proggr] (exec) {Hardware};\\
    };

    \node[fill=white,right of=exec,node distance=3cm] (grap) {\scriptsize \textcolor{vert}{Clusters}};
    \node[fill=white,right of=grap,node distance=2cm] (mc) {\scriptsize \textcolor{vert}{Multi-cores}};
    \node[fill=white,right of=mc,node distance=2cm] (gpu) {\scriptsize \textcolor{vert}{GPGPUs}};
    \node[fill=white,right of=gpu,node distance=2cm] (many) {\scriptsize \textcolor{vert}{Many-cores ...}};

    \node[fill=white,right of=para,node distance=3.5cm] (donn) {\scriptsize \textcolor{blue}{Data parallelism}};
    \node[fill=white,right of=donn,node distance=2.7cm] (tach) {\scriptsize \textcolor{blue}{Task parallelism}};
    \node[fill=white,right of=tach,node distance=2.7cm] (mess) {\scriptsize \textcolor{blue}{Message passing ...}};

    \node[fill=white,right of=mod,node distance=3cm] (bsp) {\scriptsize \textcolor{red}{BSP}};
    \node[fill=white,right of=bsp,node distance=1.5cm] (mpi) {\scriptsize \textcolor{red}{MPI}};
    \node[fill=white,right of=mpi,node distance=1.5cm] (openmp) {\scriptsize \textcolor{red}{OpenMP}};
    \node[fill=white,right of=openmp,node distance=1.5cm] (cuda) {\scriptsize \textcolor{red}{Cuda}};
    \node[fill=white,right of=cuda,node distance=1.5cm] (opencl) {\scriptsize \textcolor{red}{OpenCL ...}};

    \begin{scope} [every path/.style=line]
      \path (para) -- (mod);
      \path (mod) -- (exec);

      \path (grap) edge[->,thin, bend right=50] (mc);
      \path (grap) edge[->,thin, bend right=50] (gpu);
      \path (grap) edge[->,thin, bend right=50] (many);
      \path (mpi) edge[->,thin, bend right=50] (openmp);
      \path (mpi) edge[->,thin, bend right=50] (cuda);
      \path (bsp) edge[->,thin, bend right=50] (mpi);

      \path (mpi) -- (donn);
      \path (mpi) -- (tach);
      \path (mpi) -- (mess);
      \path (bsp) -- (donn);
      \path (bsp) -- (mess);
      \path (openmp) -- (tach);
      \path (openmp) -- (donn);
      \path (cuda) -- (tach);
      \path (opencl) -- (tach);
      \path (opencl) -- (mess);

      \path (mpi) -- (grap);
      \path (mpi) -- (mc);
      \path (mpi) -- (many);
      \path (bsp) -- (grap);
      \path (bsp) -- (mc);
      \path (cuda) -- (gpu);
      \path (openmp) -- (many);
      \path (openmp) -- (mc);
      \path (opencl) -- (gpu);
      \path (opencl) -- (mc);
      \path (opencl) -- (many);
      \path (opencl) -- (grap);
    \end{scope}

  \end{tikzpicture}

{\small\textbf{ETP4HPC}: programming models for non experts to reach exascale!}

}
  \end{center}
\end{frame}
%----------
%-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{"Please Help"}
%% %implicitly parallel (one sep. of concerns) / specific / performance
%% \begin{block}{HPC}
%% \begin{itemize}
%% \item Heterogenous and complex parallel hardware
%% \item Many programming models to combine
%% \item Expert programming
%% \end{itemize}
%% \end{block}

%% \begin{block}{Users}
%% Weather, climat, physics, biology etc.
%% \end{block}

%% \begin{center}
%% Separation of concerns between domain/parallelization and optimization\\
%% Domain specific languages (DSL)\\
%% \medskip
%% \resizebox{2cm}{!}{\includegraphics{images/scientist.jpg}}
%% \end{center}
%% \end{frame}
%-------------------------------------------------------------
\begin{frame}
  \frametitle{Domain Specific Languages} % HPC DSL
  \begin{block}{Advantages}
    \begin{itemize}
    \item Easy language for end users (can be application specific)
    \item Separation of concerns (domain/implementation)
    \item Implicit parallelization and optimizations
    \end{itemize}
  \end{block}
  \begin{alertblock}{Limitations}
    \begin{itemize}
    \item Difficulties deported to the DSL designer and implementer
      \begin{itemize}
      \item Low level high performance programming
      \item Maintainability and portability
      \end{itemize}
    \item Difficult to combine DSLs (interoperability)
      \begin{itemize}
      \item exascale applications = many specific domains and interactions
      \end{itemize}
    \end{itemize}
  \end{alertblock}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
  \frametitle{Component Models} % components
  \begin{block}{Overview}
    \begin{itemize}
    \item Technology advocating composition rather than development
      \begin{itemize}
      \item Old idea (late 60') but major development in the 90'
      \item After Object and before Model Driven Engineering (MDE)
      \end{itemize}
    \item Component = black box with well defined interactions
      \begin{itemize}
      \item Provides / Requires
      \end{itemize}
    \item Application = Assembly of components
    \end{itemize}
  \end{block}
  \begin{block}{Benefits}
    \begin{itemize}
      \item Application structure well defined
    \item Code-reuse and productivity
    \item Maintainability through separation of concerns
    \end{itemize}
  \end{block}
\vspace{-4cm}\hspace{8.15cm}\resizebox{2.5cm}{!}{\includegraphics{images/assembly.png}}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{L2C: A HPC Component Model}
\begin{block}{A minimalist component model for HPC}
  \begin{itemize}
  \item Component creation/deletion, configuration, and connection
  \item No L2C code between components \@ runtime
    \begin{itemize}
      \item No language interoperability provided
    \end{itemize}
  \item  Support of native interactions
    \begin{itemize}
    \item Use/Provide: C++, CORBA, [ FORTRAN (2008)]
    \item MPI communicator
    \end{itemize}
  \item \url{hlcm.gforge.inria.fr}
  \end{itemize}
\end{block}
\begin{alertblock}{Ongoing extensions}
  \begin{itemize}
  \item Concurrent reconfiguration
    \begin{itemize}
    \item DirectL2C
    \end{itemize}
  \item Task support
    \begin{itemize}
    \item {\color{red} Cf next talk on Comet!}
    \end{itemize}
  \end{itemize}
\end{alertblock}
\vspace{-4.2cm}\hspace{6.8cm}\resizebox{4cm}{!}{\includegraphics{images/jacobi-2X2.png}}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{This talk}
%implicitly parallel (one sep. of concerns) / specific / performance
\begin{block}{From a DSL: Multi Stencil Language}
\begin{itemize}
\item Descriptive language,
\item for Multi-Stencil simulations,
\item without numerical code.
\end{itemize}
\end{block}

\begin{block}{To a generated component assembly}
\begin{itemize}
\item with automatic synchronizations for data and task parallelism,
\item with empty functions to fill (components),
\item with good performance.
\end{itemize}
\end{block}

Separation of concerns between domain/implementation/parallelization

\end{frame}
%-------------------------------------------------------------
%% %-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{Motivation} % components
%% What if a DSL produces a component-based runtime ?

%% \begin{itemize}
%% \item Is it feasible ?
%% \item Is it efficient ?
%% \item \textcolor{gray}{Does it improve issues of DSLs ?}
%% \begin{itemize}
%% \item \textcolor{gray}{maintainability}
%% \item \textcolor{gray}{portability}
%% \item \textcolor{gray}{productivity}
%% \end{itemize}
%% \end{itemize}

%% Let's take a useful example: \textit{the Multi-Stencil Language }!
%% \end{frame}
%% %-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{How to improve it?} % HPC DSL
%% \begin{itemize}
%% \item Choose the good abstraction level to
%% \begin{itemize}
%% \item stay efficient
%% \item be used in many domains (find common meta-domains)
%% \end{itemize}
%% \item Choose good programming models for
%% \begin{itemize}
%% \item the compiler programming
%% \item the back-end
%% \end{itemize}
%% \item Try to reuse existing DSLs (compilation process or back-end codes)
%% \end{itemize}
%% % \textit{Delite (Scala)\\
%% % Composition and Reuse with Compiled Domain-Specific Languages\\
%% % ECOOP'13}
%% \end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{The Multi-Stencil Language}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: the mesh}
\begin{center}
\resizebox{6cm}{!}{\includegraphics{images/mesh.pdf}}
\end{center}
\begin{columns}
\begin{column}{6.5cm}
%    \fontsize{11}{11}\selectfont
  \begin{lstlisting}[frame=single]
mesh : cart
mesh entities : cell, edgex
computation domains :
  d1 in cell
  d2 in edgex
stencil shapes : 
  ncc from cell to cell
  nce from cell to edgex
  nec from edgex to cell
\end{lstlisting}
\end{column}
\begin{column}{5.5cm}
\begin{itemize}
\item A Cartesian mesh is used
\item Two kinds of mesh entities are declared onto it
\item Two computations domains, one for each entity type
\item Three stencil shapes (neighborhood from mesh entity to mesh entity)
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: quantity}
\begin{columns}
\begin{column}{5cm}
\begin{lstlisting}[frame=single]
quantity :
  A,cell
  B,cell
  C,edgex
  D,cell
  E,cell
  F,cell
  G,cell
  H,edgex
  I,cell
  J,cell
scalar : mu, tau
\end{lstlisting}
\end{column}
\begin{column}{6cm}
\begin{itemize}
\item A is a quantity applied onto cell
\item C is a quantity applied onto edgex
\item mu and tau are scalar values
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{The MSL language: time and computations}
\begin{columns}
\begin{column}{6cm}
\begin{lstlisting}[frame=single]
time : 500
computations :
  B[d1] = k0({tau},{A})
  C[d2] = k1({},{B[nce]})
  D[d1] = k2({},{C})
  E[d1] = k3({},{C})
  F[d1] = k4({},{D,C[nec]})
  G[d1] = k0({mu,tau},{E})
  H[d2] = k6({},{F})
  I[d1] = k7({},{G,H})
  J[d1] = k8({mu},{I[ncc]})
\end{lstlisting}
\end{column}
\begin{column}{5cm}
\begin{itemize}
\item The time loop is composed of 500 iterations.
\item J is written
\item onto the computation domain d1,
\item by the computation k8,
\item which read the scalar mu
\item and the quantity I,
\item which is accessed with the neighborhood ncc.
\end{itemize}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------

%% %-------------------------------------------------------------
%% \section{Multi-Stencil Language : Yet another DSL for stencils !}
%% %-------------------------------------------------------------
%% \begin{frame}
%% \frametitle{Multi-Stencil application}
%% \begin{center}
%% \resizebox{\textwidth}{!}{%
%% \begin{tikzpicture}
%% \def \n {5}
%% \def \radius {3cm}
%% \def \margin {8} % margin in angles, depends on the radius

%%   \node[purp2,align=center] (edp) at ({360/\n * (2 - 1)}:\radius) {Partial Differential Equations\\\scriptsize $\frac{\partial u(x,y,t)}{\partial t} = \frac{\partial^2 u(x,y,t)}{\partial x^2} + \frac{\partial^2 u(x,y,t)}{\partial y^2}$};
%%   \node[progm,align=center,right of=edp,node distance=6cm] (cond) {\scriptsize + specific behavior for\\\scriptsize boundary conditions};
%%   \path[line,dashed] (edp) -- (cond);
%%   \node[purp,align=center] (maillage) at ({360/\n * (1 - 1)}:\radius) {Mesh\\Time iterations};
%%   \path[->] (maillage) edge[<-,thin,shorten <=2pt,shorten >=2pt, bend right=40]  node [fill=white,align=center] {\scriptsize Time and space\\\scriptsize discretization} (edp.east);
%%   \path[] (edp.west) edge[thin,shorten <=2pt,shorten >=2pt, bend right=90]  node [draw,circle,fill=white] (plus) {\scriptsize +} (maillage.west);
%%   \node[purp,align=center, below of=plus, node distance=3.5cm] (sche) {Explicit\\numerical schemes\\\textcolor{red}{stencils}};
%%   \node[progm, text width=11em,align=center,right of=sche,node distance=5.5cm] (expl) {\scriptsize Iterative neighboorhood approximation of the phenomena};
%%   \path[line,dashed] (sche) -- (expl);
%%   \path[line] (plus) -- node [fill=white,align=center] {\scriptsize Numerical methods\\\scriptsize Finite difference/volume/elment} (sche);
%% \end{tikzpicture}
%% }
%% \end{center}
%% \end{frame}
%% %----------------
%% \begin{frame}
%% \frametitle{Multi-Stencil Language : Yet another DSL for stencils !}
%% \begin{block}{Inputs}
%% follow a descriptive grammar to simply declare
%% \begin{itemize}
%% \item the sequential order of computations
%% \item what is read and written by each computation
%% \item with or without a neighborhood access (stencil)
%% \end{itemize}
%% \end{block}
%% \begin{block}{Outputs}
%% \begin{itemize}
%% \item parallel pattern of the multi-stencil application
%% \item hybrid parallelism included: data and task parallelism
%% \item choice of parallel languages (MPI, OpenMP 3.x/4.x, SkelGIS etc.)
%% \item end-users have to fill computation functions into this pattern
%% \end{itemize}
%% \end{block}
%% \end{frame}
%% %----------------
%% \begin{frame}
%% \frametitle{Multi-Stencil Language : Yet another DSL for stencils !}
%% \begin{itemize}
%% \item Choosen abstraction level
%% \begin{itemize}
%% \item mesh-agnostic description language
%% \begin{itemize}
%% \item should be usable for unstructured, Cartesian, curvilinear meshes
%% \end{itemize}
%% \item description concerns splited from implementation concerns
%% \begin{itemize}
%% \item usable with many different parallel libraries or languages back-ends
%% \end{itemize}
%% \end{itemize}
%% \item Choosen programming models for back-end
%% \begin{itemize}
%% \item Component programming model (maintainability, portability and composability)
%% \item Task scheduling model (portability and efficiency)
%% \end{itemize}
%% \item Reuse of existing languages and DSLs
%% \begin{itemize}
%% \item Use of a Distributed Data Structure DSeL (C++,MPI) : SkelGIS
%% \item MPI and OpenMP (3.x and 4.x)
%% \end{itemize}
%% \end{itemize}
%% \end{frame}
%% %-------------------------------------------------------------

%-------------------------------------------------------------
\section{From DSL to Component Assembly}
%-------------------------------------------------------------
%----------------
\begin{frame}
\frametitle{Data and Task parallelism}
\begin{block}{Synchronizations}
When a computation read a data, using a stencil shape, that has been written by a previous computation.
\end{block}

$\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$\\
$\hookrightarrow [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$

\medskip

\begin{block}{Dependency graph}
%\begin{enumerate}
Node: a computation or a synchronization\\
Edge: a direct data dependency (Read / Write)
\end{block}

\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node (c0) at (0,0) {$c_0$};
   \node[right=1 of c0] (sy1) {$sync_1$};
   \node[right=1 of sy1] (c1) {$c_1$};
   \node[above right=1 of c1] (c2) {$c_2$};
   \node[below right=1 of c1] (c3) {$c_3$};
   \node[right=1 of c2] (sy4) {$sync_4$};
   \node[right=1 of c3] (c5) {$c_5$};
   \node[right=1 of sy4] (c4) {$c_4$};
   \node[right=1 of c4] (c6) {$c_6$};
   \node[below right=1 of c6] (c7) {$c_7$};
   \node[right=1 of c7] (sy8) {$sync_8$};
   \node[right=1 of sy8] (c8) {$c_8$};

  \path[->]
    (c0) edge node {} (sy1)
    (sy1) edge node {} (c1)
    (c1)  edge node {} (c2)
          edge node {} (c3)
    (c2) edge node {} (sy4)
    (sy4) edge node {} (c4)
    (c4) edge node {} (c6)
    (c3) edge node {} (c5)
    (c5) edge node {} (c7)
    (c6) edge node {} (c7)
    (c7) edge node {} (sy8)
    (sy8) edge node {} (c8);
\end{tikzpicture}
\end{center}

\begin{center}
\textcolor{red}{Static} or \textcolor{red}{dynamic} scheduling ?
\end{center}
\end{frame}
%----------------
%----------------
\begin{frame}
\frametitle{Series-Parallel Tree}
\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
\normalsize
\begin{columns}
\begin{column}{.48\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

   \node[] (c0)    at (-3,1) {$c_0$};
   \node[] (star1) at (-2,1) {$sync_1$};
   \node[] (c1)    at (-1,1) {$c_1$};
   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
   \node[] (c7)    at ( 1,1) {$c_7$};
   \node[] (star8) at ( 2,1) {$sync_8$};
   \node[] (c8)    at ( 3,1) {$c_8$};

   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
   \node[] (s2)    at (1,2) {$\mathcal{S}$};

   \node[] (c2)    at (-3,3) {$c_2$};
   \node[] (star4) at (-2,3) {$sync_4$};
   \node[] (c4)    at (-1,3) {$c_4$};
   \node[] (c6)    at ( 0,3) {$c_6$};
   \node[] (c3)    at ( 1,3) {$c_3$};
   \node[] (c5)    at ( 2,3) {$c_5$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3)
         edge node {} (c5);
  \end{tikzpicture}
  }
\end{center}
\end{column}
\begin{column}{.52\textwidth}

\end{column}
\end{columns}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Series-Parallel Tree}
\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
\normalsize
\begin{columns}
\begin{column}{.48\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

   \node[] (c0)    at (-3,1) {$c_0$};
   \node[] (star1) at (-2,1) {$sync_1$};
   \node[] (c1)    at (-1,1) {$c_1$};
   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
   \node[] (c7)    at ( 1,1) {$c_7$};
   \node[] (star8) at ( 2,1) {$sync_8$};
   \node[] (c8)    at ( 3,1) {$c_8$};

   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
   \node[] (s2)    at (1,2) {$\mathcal{S}$};

   \node[] (c2)    at (-3,3) {$c_2$};
   \node[] (star4) at (-2,3) {$sync_4$};
   \node[] (c4)    at (-1,3) {$c_4$};
   \node[] (c6)    at ( 0,3) {$c_6$};
   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
   %\node[] (c5)    at ( 2,3) {$c_5$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3);
  \end{tikzpicture}
  }
\end{center}
\end{column}
\begin{column}{.52\textwidth}
\end{column}
\end{columns}
Loop fusion optimization possible
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Series-Parallel Tree}
\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
\normalsize
\begin{columns}
\begin{column}{.48\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

   \node[] (c0)    at (-3,1) {$c_0$};
   \node[] (star1) at (-2,1) {$sync_1$};
   \node[] (c1)    at (-1,1) {$c_1$};
   \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
   \node[] (c7)    at ( 1,1) {$c_7$};
   \node[] (star8) at ( 2,1) {$sync_8$};
   \node[] (c8)    at ( 3,1) {$c_8$};

   \node[] (s1)    at (-1,2) {$\mathcal{S}$};
   \node[] (s2)    at (1,2) {$\mathcal{S}$};

   \node[] (c2)    at (-3,3) {$c_2$};
   \node[] (star4) at (-2,3) {$sync_4$};
   \node[] (c4)    at (-1,3) {$c_4$};
   \node[] (c6)    at ( 0,3) {$c_6$};
   \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
   %\node[] (c5)    at ( 2,3) {$c_5$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3);
  \end{tikzpicture}
  }
\end{center}
\end{column}
\begin{column}{.52\textwidth}
\begin{block}{Specific components}
\begin{itemize}
\item $SEQ$ to directly replace $\mathcal{S}$ nodes
\item $PAR$ to directly replace $\mathcal{P}$ nodes
\item $SYNC$ for synchronizations
\item $K$ for computation kernels
\end{itemize}
\end{block}
\end{column}
\end{columns}
Loop fusion optimization possible
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{MSL to Component-based runtime}
\begin{block}{Ready-to-fill parallel pattern}
\begin{itemize}
\item Data parallelism
\begin{itemize}
\item External distributed data structure
\item Automatic detection of synchronizations
\end{itemize}
\item Task parallelism (mid-grain)
\begin{itemize}
\item Static scheduling at the assembly level 
\item Dynamic scheduling within a driver component
  \begin{itemize}
  \item OpenMP task
  \end{itemize}
\end{itemize}
\end{itemize}
\end{block}
The fine grain task parallelism is left to other languages: 
\begin{itemize}
\item OpenMP in the kernels
\item Kernels generated by stencil compilers
  \begin{itemize}
  \item Pochoir, PATUS, Liszt etc.
  \end{itemize}
\end{itemize}
\end{frame}
%----------------
%-------------------------------------------------------------
\begin{frame}
\frametitle{Component Assembly Runtime Template}
A data parallel/task parallel component assembly.\vspace{-0.2em}

\begin{center}
\begin{tikzpicture}[remember picture]
  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
    \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
     \node[component,solid,draw=black] (D) at (0,0) {$Driver$};
     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
     \node[solid] (Ds) at (-1.5,0) {start};
     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};

     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
     \node[solid,left=0.8cm of T] (tt) {$T$};

     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
     \node[component,solid,right=2cm of Cp] (C) {$Scheduler$};
     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
     \node[use,scale=0.5,solid,right=1.5cm of C] (Cu2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Cu2] (Cu3) {};
     \node[use,scale=0.5,solid,below=0.8cm of Cu3] (Cu4) {};
     \node[solid,right=0.2cm of Cu] (star) {$*$};
     %\node[solid,right=1.5cm of C] (gamma) {$\Gamma$};

     \node[solid,below=1cm of C] (calc) {$\Gamma_{sync}$, $\Gamma_{dep}$};

     \node[provide,scale=0.5,solid,right=0.15 of Cu2] (k0p) {};
     \node[component,solid,right=1.5cm of k0p] (k0) {$k_0$};

     \node[provide,scale=0.5,solid,right=0.15 of Cu3] (k0sp) {};
     \node[component,solid,right=1.5cm of k0sp] (k0s) {$k_{0;1}^{sync}$};

     \node[provide,scale=0.5,solid,right=0.15 of Cu4] (kotherp) {};
     \node[component,solid,right=1.5cm of kotherp] (kother) {$\dots$};

     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Quantity$};
     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
     \node[solid,left=1.5cm of Data] (delta) {$\Delta,\mathcal{D}$};
     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};

     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
     \node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};

    \path[-]
      (Dp) edge [solid] node {} (D)
      (D) edge [solid] node {} (Du1)
          edge [solid] node {} (Du2)
          edge [solid] node {} (Du3)
      (DDSp1) edge [solid] node {} (DDS)
      (Tp) edge [solid] node {} (T)
      (T)  edge [solid] node {} (Tu)
      (Cp) edge [solid] node {} (C)
      (C) edge [solid] node {} (Cu)
      (Datap2) edge [solid] node {} (Data)
      (Datap1) edge [solid] node {} (Data)
      (Data) edge [solid] node {} (Datau)
      (DDSp2) edge [solid] node {} (DDS)
      (C)  edge [solid] node {} (Cu2)
          edge [solid] node {} (Cu3)
          edge [solid] node {} (Cu4)
      (Cu2)  edge [solid] node {} (k0p)
      (k0p)  edge [solid] node {} (k0)
      (Cu3)  edge [solid] node {} (k0sp)
      (k0sp)  edge [solid] node {} (k0s)
      (Cu4)  edge [solid] node {} (kotherp)
      (kotherp)  edge [solid] node {} (kother);
    \end{tikzpicture}
    };
    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=4cm of Data] (mpidata) {};
    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
    \node[solid,above=3.2cm of proc.west,anchor=west] (core) {\textit{\tiny{Applied on each resource}}};
    \path[-]
      (Data.east) edge node {} (mpidata);
  \end{tikzpicture}
\end{center}
{\fontsize{8}{8}\selectfont DDS: Distributed Data Structure (using SkelGIS)\par}
\end{frame}
%----------------
\begin{frame}
\frametitle{Component-based runtime: Static Scheduling}

\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
     \node[component,solid] (D) at (0,0) {$Driver$};
     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
     %\node[solid] (Ds) at (-1.5,0) {start};
     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};

     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};
     %\node[solid,left=0.8cm of T] (tt) {$T$};

     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Cp) {};
     \node[component,dashed,right=2cm of Cp] (C) {$Computations$};
     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};
     \node[solid,right=0.2cm of Cu] (star) {$*$};
     %\node[solid,right=1.5cm of C] (gamma) {$\Gamma$};

     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};
     %\node[solid,left=1cm of Data] (delta) {$\Delta,\mathcal{D}$};
     %\node[mpi,scale=0.8,rounded corners=0pt,solid,right=1cm of Data] (mpidata) {};
     %\node[solid,above=0.2cm of mpidata] (star2) {$*$};

     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};
     %\node[solid,right=1.2cm of DDS] (m) {$\mathcal{M},\mathcal{E}$};

    \path[-]
      (Dp) edge [solid] node {} (D)
      (D) edge [solid] node {} (Du1)
          edge [solid] node {} (Du2)
          edge [solid] node {} (Du3)
      (DDSp1) edge [solid] node {} (DDS)
      (Tp) edge [solid] node {} (T)
      (T)  edge [solid] node {} (Tu)
      (Cp) edge [solid] node {} (C)
      (C) edge [solid] node {} (Cu)
      (Datap2) edge [solid] node {} (Data)
      (Datap1) edge [solid] node {} (Data)
      (Data) edge [solid] node {} (Datau)
      (DDSp2) edge [solid] node {} (DDS);
    %\end{tikzpicture}
    %}
% \end{column}
% \begin{column}{.70\textwidth}
% \begin{center}
% \resizebox{.80\textwidth}{!}{%
% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   %seq0
   \node[component,right=3.5cm of C] (seq0) {$SEQ$};
   \node[provide] (seq0p) [left = 0.8cm of seq0] {};
   \node[use] (seq0u) [right = 1cm of seq0] {$m$};
   %c0
   \node[component] (c0) [right = 2cm of seq0u] {$K(c_0)$};
   \node[use] (c0u) [right = 0.8cm of c0] {};
   \node[right=0.2 of c0u] (star) {*};
   %sync0
   \node[component] (sync0) [below = 1cm of c0] {$SYNC$};
   \node[use] (sync0u) [right = 1cm of sync0] {$m$};
   %c1
   \node[component] (c1) [below = 0.8cm of sync0] {$K(c_1)$};
   \node[use] (c1u) [right = 0.8cm of c1] {};
   \node[right=0.2 of c1u] (star) {*};
   %par0
   \node[component] (par0) [below = 0.8cm of c1] {$PAR$};
   \node[use] (par0u) [right = 1cm of par0] {$m$};
  %seq1
   \node[component] (seq1) [right = 1cm of sync0u] {$SEQ$};
   \node[use] (seq1u) [right = 1cm of seq1] {$m$};
  %c2
   \node[component] (c2) [right = 4cm of c0] {$K(c_2)$};
   \node[use] (c2u) [right = 0.8cm of c2] {};
   \node[right=0.2 of c2u] (star) {*};
  %sync1
   \node[component] (sync1) [below = 0.8cm of c2] {$SYNC$};
   \node[use] (sync1u) [right = 1cm of sync1] {$m$};
  %c4
   \node[component] (c4) [below = 0.8cm of sync1] {$K(c_4)$};
   \node[use] (c4u) [right = 0.8cm of c4] {};
   \node[right=0.2 of c4u] (star) {*};
  %c6
   \node[component] (c6) [below = 0.8cm of c4] {$K(c_6)$};
   \node[use] (c6u) [right = 0.8cm of c6] {};
   \node[right=0.2 of c6u] (star) {*};
  %seq2
   \node[component] (seq2) [below = 3.2cm of seq1] {$SEQ$};
   \node[use] (seq2u) [right = 1cm of seq2] {$m$};
  %c3
   \node[component] (c3) [right = 1cm of seq2u] {$K(c_3)$};
   \node[use] (c3u) [right = 0.8cm of c3] {};
   \node[right=0.2 of c3u] (star) {*};
  %c5
   \node[component] (c5) [below = 0.8cm of c3] {$K(c_5)$};
   \node[use] (c5u) [right = 0.8cm of c5] {};
   \node[right=0.2 of c5u] (star) {*};
  %c7
   \node[component] (c7) [below = 0.8cm of par0] {$K(c_7)$};
   \node[use] (c7u) [right = 0.8cm of c7] {};
   \node[right=0.2 of c7u] (star) {*};
  %sync2
   \node[component] (sync2) [below = 0.8cm of c7] {$SYNC$};
   \node[use] (sync2u) [right = 1cm of sync2] {$m$};
  %c8
   \node[component] (c8) [below = 0.8cm of sync2] {$K(c_8)$};
   \node[use] (c8u) [right = 0.8cm of c8] {};
   \node[right=0.2 of c8u] (star) {*};

   \path[-]
   %seq0
    (seq0) edge node {} (seq0u)
    (seq0p) edge node {} (seq0)
   %c0
    (seq0u) edge node {} (c0)
    (c0) edge node {} (c0u)
    %sync0
    (seq0u) edge node {} (sync0.west)
    (sync0) edge node {} (sync0u)
    %c1
    (seq0u) edge node {} (c1)
    (c1) edge node {} (c1u)
   %par0
    (seq0u) edge node {} (par0.west)
    (par0) edge node {} (par0u)
  %seq1
    (par0u) edge node {} (seq1.west)
    (seq1) edge node {} (seq1u)
  %c2
    (seq1u) edge node {} (c2.west)
    (c2) edge node {} (c2u)
  %sync0
    (seq1u) edge node {} (sync1.west)
    (sync1) edge node {} (sync1u)
  %c4
    (seq1u) edge node {} (c4.west)
    (c4) edge node {} (c4u)
  %c6
    (seq1u) edge node {} (c6.west)
    (c6) edge node {} (c6u)
  %seq2
    (par0u) edge node {} (seq2.west)
    (seq2) edge node {} (seq2u)
  %c3
    (seq2u) edge node {} (c3)
    (c3) edge node {} (c3u)
  %c6
    (seq2u) edge node {} (c5.west)
    (c5) edge node {} (c5u)
  %c6
    (seq0u) edge node {} (c7.west)
    (c7) edge node {} (c7u)
  %sync2
    (seq0u) edge node {} (sync2.west)
    (sync2) edge node {} (sync2u)
  %c8
    (seq0u) edge [bend right] node {} (c8.west)
    (c8) edge node {} (c8u)
   ;

   \path[-]
    (C) edge[dashed] node {} (seq0);
\end{tikzpicture}
}
\end{center}
% \end{column}
% \end{columns}
\end{frame}
%----------------
\begin{frame}
  \frametitle{Component-based runtime: Dynamic Scheduling}
\begin{center}
  \begin{itemize}
  \item \emph{mpiOmpDyn}
    \begin{itemize}
    \item \emph{DDS} and \emph{Quantity}: data parallelization.
    \item \emph{Scheduler}: DAG encoded into OpenMP tasks
    \end{itemize}
  \end{itemize}
  \medskip
  \begin{tikzpicture}[remember picture]
  \node[rectangle,rounded corners=3pt,thick,inner sep=5pt,draw=black,dotted] (proc) {
    \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
     \node[component,solid] (D) at (0,0) {$Driver$};
     \node[provide,scale=0.5,solid] (Dp) at (-1,0) {};
     \node[solid] (Ds) at (-1.5,0) {start};
     \node[use,scale=0.5,solid,right=1.5cm of D] (Du1) {};
     \node[use,scale=0.5,solid,below=1.75cm of D] (Du2) {};
     \node[use,scale=0.5,solid,below=0.8cm of Du1] (Du3) {$m$};

     \node[provide,scale=0.5,solid,below=0.15 of Du2] (Tp) {};
     \node[component,solid,below=1.6cm of Tp] (T) {$Time$};
     \node[use,scale=0.5,solid,right=1cm of T] (Tu) {};

     \node[provide,scale=0.5,solid,right=0.15 of Tu] (Sp) {};
     \node[component,solid,right=1.5cm of Sp] (S) {$Scheduling$};
     \node[use,scale=0.5,solid,right=1.5cm of S] (Su) {};

     \node[provide,scale=0.5,solid,right=0.15 of Su] (Cp) {};
     \node[component,solid,double,right=2cm of Cp] (C) {$Kernel$};
     \node[use,scale=0.5,solid,above=0.8cm of C] (Cu) {};

     \node[provide,scale=0.5,solid,below right=0.2 of Du3] (Datap1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Cu] (Datap2) {};
     \node[component,solid,double,above=0.8cm of Datap2] (Data) {$Data$};
     \node[use,scale=0.5,solid,above=0.8cm of Data] (Datau) {};

     \node[provide,scale=0.5,solid,right=0.15 of Du1] (DDSp1) {};
     \node[provide,scale=0.5,solid,above=0.15 of Datau] (DDSp2) {};
     \node[component,solid,above=0.8cm of DDSp2] (DDS) {$DDS$};

    \path[-]
      (Dp) edge [solid] node {} (D)
      (D) edge [solid] node {} (Du1)
          edge [solid] node {} (Du2)
          edge [solid] node {} (Du3)
      (DDSp1) edge [solid] node {} (DDS)
      (Tp) edge [solid] node {} (T)
      (T)  edge [solid] node {} (Tu)
      (Sp) edge [solid] node {} (S)
      (S) edge [solid] node {} (Su)
      (Datap2) edge [solid] node {} (Data)
      (Datap1) edge [solid] node {} (Data)
      (Data) edge [solid] node {} (Datau)
      (C) edge [solid] node {} (Cu)
      	  edge [solid] node {} (Cp)
      (DDSp2) edge [solid] node {} (DDS);
    \end{tikzpicture}
    };
    \node[mpi,scale=0.8,rounded corners=0pt,solid,right=0.5cm of proc] (mpidata) {};
    \node[solid,above=0.1cm of mpidata] (star2) {$*$};
    \node[solid,below=2.5cm of proc.west,anchor=west] (core) {\textit{\tiny{Duplicated on each processor/core}}};
    \path[-]
      (Data.east) edge node {} (mpidata);
  \end{tikzpicture}
\end{center}

\end{frame}

%-------------------------------------------------------------
\section{Evaluations}
%----------------
\begin{frame}
\frametitle{Evaluation Setup}
\begin{block}{Application}
\begin{itemize}
\item FullSWOF2D: developed at the MAPMO, University of Orléans
\item Solve the shallow-water equations using a finite volume method
\item Cartesian mesh
\item 3 mesh entities, 7 computation domains, 48 quantities
\item 98 computations (32 stencils, 66 local kernels)
\end{itemize}
\end{block}
\begin{block}{Machines}
\begin{itemize}
\item Thin nodes TGCC Curie
\item 2 CPU, 8 cores, Sandy Bridge EP (E5-2680) 2.7 GHz, 64 GB
\item OpenMPI and gcc 4.9
\end{itemize}
\end{block}
\end{frame}
%----------------
%-------------------------------------------------------------
\begin{frame}
\frametitle{Data Parallelization: Weak and Strong Scaling}
\begin{columns}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/median_weak.pdf}}
\begin{center}
\vspace{-1em}
  Weak Scaling\\
  400x400 domain/core
\end{center}
\end{column}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/median_strong.pdf}}
\begin{center}
\vspace{-1em}
  Strong Scaling\\
  10k x 10k domain
\end{center}
\end{column}
\end{columns}
\bigskip
\begin{center}
  \textbf{No overhead introduced by components}
\end{center}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Data Parallelization to Hybrid Parallelization: Strong Scaling}
\begin{columns}
\begin{column}{6cm}
  \resizebox{6cm}{!}{\includegraphics{images/times.pdf}}
  \begin{center}
{\vspace{-1em}\small    Data Parallelization    }
  \end{center}
\end{column}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/base_close_median.pdf}}
\begin{center}
{\vspace{-1em}\small  Data and Hybrid Parallelizations}
\end{center}
\end{column}
\end{columns}
\medskip
\begin{center}
{\small
  \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline 
   Parallelism Level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
   \hline
   Frequency & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
   \hline
 \end{tabular}
  \\
  \medskip
 Task graph analysis
}
\end{center}
Mesh: 500x500
\end{frame}
%-------------------------------------------------------------

\begin{frame}
  \frametitle{Dynamic Scheduling and Energy Evaluations}

  Application versions:
  \small
  \centering
\begin{tabular}{|c||c|c|c|}
  \hline
  Variant Name  & DDS / Quantity & Kernels & Scheduler \\
  \hline
  \hline
  mpiBase        & MPI & Sequential & Sequential \\
  \hline
  mpiOmpFor      & MPI & Parallel loops & Sequential \\
  \hline
  mpiOmpForkJoin & MPI & Sequential & SPT \\
  \hline
  mpiOmpDyn      & MPI & Sequential & OpenMP Tasks \\
  \hline
\end{tabular}

\end{frame}

%-------------------------------------------------------------

  \begin{frame}
\frametitle{Energy Usage on Grid'5000}
\begin{columns}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/4000_100_4node_med_time_maxWatt_cloud.pdf}}
    \begin{center}
{\vspace{-1em}\small Time vs Max Watt on 4 nodes}
  \end{center}
\end{column}
\begin{column}{6cm}
  \resizebox{6cm}{!}{\includegraphics{images/4000_100_4node_med_time_avrgWatt.pdf}}
\begin{center}
{\vspace{-1em}\small Time vs Average Watt on 4 nodes}
\end{center}
\end{column}
\end{columns}
\bigskip
{\small
\begin{block}{Parameters:}
\begin{itemize}
\vspace{-0.7em}\item Domain: 4000 x 4000
\item \#Iteration: 100
\end{itemize}
\end{block}
}
  \end{frame}
%-------------------------------------------------------------
  \begin{frame}

\frametitle{Energy Usage on Curie}
\begin{columns}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/Curie_20k_10k_time_avrgWatt_2048.pdf}}
    \begin{center}
{\vspace{-1em}\small Time vs Max Watt on 2048 cores}
  \end{center}
\end{column}
\begin{column}{6cm}
  \resizebox{6cm}{!}{\includegraphics{images/Curie_84k_500_time_avrgWatt.pdf}}
\begin{center}
{\vspace{-1em}\small Time vs Average Watt on 2048 cores}
\end{center}
\end{column}
\end{columns}
\bigskip
{\small
\begin{block}{Parameters:}
\begin{itemize}
\vspace{-0.7em}\item Domain: 20,000 x 20,000
\item \#Iteration: 10,000
\item One measure every 5 min!
\end{itemize}
\end{block}
}

\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Conclusion}
%-------------------------------------------------------------
%\subsection{Conclusion and perspectives}
\begin{frame}
\frametitle{Conclusion and Future Work}
\begin{block}{Conclusion}
\begin{itemize}
\vspace{-0.4em}\item A DSL for Multi-Stencil applications (MSL)
\item The generation of a component based runtime, including scheduling
\begin{itemize}
\item Data and task parallelism
\end{itemize}
\item Evaluation on Grid'5000 and Curie: no overhead introduced
\item Basis for ongoing energy studies
\end{itemize}
\end{block}

\begin{block}{Perspectives}
  \begin{itemize}
\vspace{-0.4em}  \item Language improvement (convergence criteria, reduction etc.)
  \item OpenMP inside kernels/components
  \item CPU+GPGPUs using stencil compilers (Pochoir, PATUS etc.)
  \item Composition of various DSLs by agreeing on component runtimes?
  \end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------

\end{document}