### 2017-02-17 ARTICLES/IJPP/articleR1/msf.tex
nouvelle partie II + ajout partie IV

\HC{under construction, present the framework}

This section first presents a background on component models and particularly on the Low Level Components. This background is needed to understand the second part of the section which presents an overview of the overall Multi-Stencil Framework (MSF).

%----------------------------------------
\subsection{Background on component models}
%----------------------------------------
Component-based software engineering (CBSE) is a domain of software engineering~\cite{Szyperski:2002:CSB:515228} which improves code re-use, separations of concerns, and thus maintainability.
An component-based application is made of a set of component instances linked together, this is also called a component \emph{assembly}.
A component is a black box that implements an independent functionality of the application, and which interacts with its environment only through well defined interfaces: its ports.
A port can, for example, specify services provided or required by the component.
With respect to high performance computing, some works have also shown
that component models can achieve the needed level of performance, and
scalability while also helping in application portability~\cite{Bernholdt01052006, bigot:inria-00388508, UCHPC2015}

Many component models exist, each of them with its own specificities.
Well known component models include, for example, the CORBA Component Model (CCM)~\cite{corba:omg06}, and the Grid Component Model (GCM)~\cite{Baude} for distributed computing, while the Common Component Architecture (CCA)~\cite{Bernholdt01052006}, and Low Level Components (\llc)~\cite{l2c} are HPC-oriented.
This work makes use of \llc for the experiments.

\llc~\cite{l2c} is a minimalist \texttt{C++} based HPC-oriented component model
where a component extends the concept of class.
The services offered by the components are specified trough $provide$ ports,
those used either by $use$ ports for a single service instance,
or $use-multiple$ ports for multiple services instances.
Services are specified as \texttt{C++} interfaces.
\llc also offers $MPI$ ports that enable components to share MPI communicators. Finally, 
components can also have attribute ports to be configured.
%
In this paper, and as illustrated in Figures~\ref{fig:ports}, a $provide$ port is represented by a white circle, a $use$ port with a black circle, a $use-multiple$ port by a black circle with a white $m$ in it. MPI port are connected with a black rectangle. A \llc-based application is a static \emph{assembly} of components instances and the connections between their ports.
Such an assembly is described in LAD, an XML dialect.

\begin{figure}[t]
\begin{center}
\subfloat[][\label{fig:2comp}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$c_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[provide,right=1.5cm of u] (p1) {};
   \node[component,right=1.5cm of p1] (C1) {$c_1$};
   \node[use,right=1.5cm of C1] (um) {$m$};

  \path[-]
    (p) edge node {$p$} (C)
    (C.east) edge node {$u$} (u)
    (C1)  edge node {$v$} (um)
    (p1) edge node {$q$} (C1);
\end{tikzpicture}
}
\\
\subfloat[][\label{fig:ass}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$c_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[provide,right=0.15 of u] (p2) {};
   \node[component,right=1.5 of p2] (C1) {$c_1$};
   \node[use,right=1.5 of C1] (um) {$m$};

  \path[-]
    (p) edge node {$p$} (C)
    (C) edge node {$u$} (u)
    (C1)  edge node {$v$} (um)
    (p2) edge node {$q$} (C1);
\end{tikzpicture}
}
\\
\subfloat[][\label{fig:mpi}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$c_2$};
   \node[provide] (p) at (-1.5,0) {};
   \node[mpi] (u) at (1.5,0) {};
   \node[component,right=1.5 of u] (C1) {$c_3$};

  \path[-]
    (p) edge node {} (C)
    (C) edge node {} (u)
    (u) edge node {} (C1);
\end{tikzpicture}
}
\caption{Example of components and their ports representation. a) Component $c_0$ has a provide port ($p$) and a use port ($u$); Component $c_1$ has also a provide port ($q$) but also a use multiple port ($v$). b) A use port is connected to a (compatible) provide port. c) Component $c_2$ and $c_3$ shares an MPI communicator.}
\label{fig:ports}
\end{center}
\end{figure}

%----------------------------------------
\subsection{Multi-Stencil Framework overview}
%----------------------------------------
The Multi-Stencil Framework helps end-users to produce high performance parallel applications for the specific case of multi-stencils. The multi-stencil domain will be formally defined in the next section. A multi-stencil program is typically a numerical simulation based on explicit numerical schemes that approximate PDEs equations with one or multiple neighborhood computations.

Figure~\ref{fig:msf} presents the overview of the Multi-Stencil Framework that will be entirely detailed throughout this paper. It is composed of four different parts presented below.

One can note in this figure that MSF is addressed to two different kinds of end-users: the \emph{numerician}, in other words the mathematician, and the \emph{developer}. Most of the time numericians do have programming knowledge, however as it is not their core domain and because of a lack of time, development is often left to engineers according to numericians needs. MSF has the interesting particularity to propose a clear separation of concerns between these two end-users by distinguishing the description language from the numerical code.

MSF also has the interesting capability to become flexible thanks to a third user possibility to interact with the framework: the High Performance Computing (HPC) specialist.

\begin{figure}[t]
\begin{center}
  \includegraphics[width=.6\textwidth]{./images/msf.pdf}
  \caption{The Multi-Stencil Framework (MSF) composed of the Multi-Stencil Language (MSL), the Generic Assembly (GA), the Multi-Stencil Compiler (MSC) and which produces a specialized assembly of components.}
  \label{fig:msf}
\end{center}
\end{figure}

\paragraph{\textbf{Multi-Stencil Language}}
The Multi-Stencil Language, or MSL, is the domain specific language proposed by the framework for the numerician. It is a descriptive language, easy to use, without any concern onto implementation details, which fits the need of a mathematician to describe the simulation. The description written with MSL can be considered as an input of the framework. MSL is described in details in Section~\ref{sect:msl}. The language is built upon the meta-model described in Section~\ref{sect:formalism}.

\paragraph{\textbf{Generic Assembly}}
In addition to the language MSL and the simulation description written by the numerician, MSF needs a generic assembly of a multi-stencil program as input. What is called a generic assembly is a component assembly for which meta-types are represented and for which some parts need to be generated or specialized. It is similar to a generic pattern or a skeleton in usual programming models such as C++ or functional models. From this generic assembly will be built the final specialized assembly of the simulation. As well as MSL, this generic assembly is described in Section~\ref{sect:msl} and is built upon the meta-model described in Section~\ref{sect:formalism}. Moreover, an example of generic assembly is used in evaluations of Section~\ref{sect:eval}.
\HC{Note idée : HPC specialist gives types for DDS and Data}

\paragraph{\textbf{Multi-Stencil Compiler}}
The core of the framework is the Multi-Stencil Compiler, or MSC. It is responsible for transforming the genric assembly into the final parallel assembly which is specific to the simulation described by the numerician. MSC is described in Sections~\ref{sect:parallelism} and~\ref{sect:msp}.

\paragraph{\textbf{Specialized assembly}}
Finally, the input of MSF is the component assembly generated by MSC. It is an instanciation and a transformation of the generic component assembly, by adding component types and by adding specific components generated by MSC. From this final component assembly which is specific to the simulation initially described by using MSL, the developer will finally write numerical codes.