### 2017-02-20 ARTICLES/IJPP/articleR1/related.tex

% orchestration of stencils, different from stencil compilers like Pochoir and PATUS which could be used inside MSF
First of all, many domain specific languages have been proposed for the optimization of a single stencil computation. Each one has its own specificities and answers to a specific stencil case or to a specific additionnal optimization. For example, Pochoir~\cite{spaaTangCKLL11} works on cache optimization techniques for stencils applied onto Cartesian meshes. On the other hand, PATUS~\cite{citeulike12258902} proposes to add a parallelization strategy grammar to its stencil language to perform an auto-tuning parallelization strategy. Finally, ExaSlang~\cite{Schmitt:2014:EDL:2691166.2691171} is specific to multigrid numerical methods etc. Thus, these stencil compilers target a different scope than the Multi-Stencil Framework presented in this paper, which actually orchestrates a parallel execution of multiple stencil codes together. Furthermore, future works could lead to a combination of these stencil compilers with MSF to build very optimized stencil kernels $K$.

% Stencil orchestration
Some solutions, closer to MSF, have also been proposed to automatically orchestrate multiple stencils and local computations in a parallel manner. Amongs them is Halide~\cite{Ragan-Kelley:2013:HLC:2491956.2462176} that proposes an optimization and parallelization of a pipeline of stencils. However, Halide is limited to Cartesian meshes and is specific to images. Liszt~\cite{DeVito:2011:LDS:2063384.2063396} and Nabla~\cite{Camier:2015:IPP:2820083.2820107} both offer solutions for the automatic parallel orchestration of stencils applied onto any kind of mesh, from Cartesian to unstructured meshes. The needed mesh can be built from a set of available symbols in the grammar of each language. Thus, these languages generalize the definition of a mesh, as it is proposed into the MSL formalism of Section~\ref{sect:formalism}. Neither Liszt or Nabla, however, handle hybrid parallelism as it is proposed by MSF.

% MSF is not a general purpose framework. It proposes a new approach to improve sep. of concerns and code-reuse in Domain Specific Languages, by combining it to component generic assemblies.
We have made the choice in MSF to propose and use the MSL Domain Specific Language to ask the numerician a sequential description of its simulation. This description, is close to a dataflow representation. However, MSL is specific to numerical simulations and proposes a mesh abstraction adapted to numerical simulations. Compared to general purpose dataflow runtimes such as Legion~\cite{bauer:legion:sc:2012}, HPX~\cite{}, PFunc~\cite{}, MSL is closer to the semantic of the domain (mesh, stencils etc.), thus easier to use for non-specialists. Moreover, MSL is very light and only descriptive as numerical codes are left to another language and another user (the developer in Figure~\ref{fig:msf}). Furthermore, these three runtimes could actually be used by MSF as a backend, instead of using SkelGIS and OpenMP.
This is due to software engineering capacities introduced by proposing a meta-model and by using component models in MSF.

% vision of the framework brings separation of concerns, code-reuse and flexibility
Actually, the vision proposed by MSF is to use components to be able to increase separation of concerns and code-reuse of the overall Framework.
%%%% sep of concerns
Separation of concerns is illustrated in Figure~\ref{fig:msf} and all along the paper. The numerician is only responsible for the description of the simulation by using MSL. A HPC specialist can adapt or propose new components for handling the distributed data structure and quantities of the simulation. MSF generates from this information the parallel orchestration of computations. Finally, after MSF compilation (MSC), the developer of numerical codes fills computation kernels by using the chosen distributed data structure. In Liszt~\cite{DeVito:2011:LDS:2063384.2063396} and Nabla~\cite{Camier:2015:IPP:2820083.2820107}, there is no such separation of concerns between the numerician and the developer. 
%%%% flexibility
Moreover, it is not possible to easily integrate a new distributed data structure in these solutions as a monolithic code is generated. By using components, doors to online reconfiguration of applications~\cite{Lanore:2015:RCM:2737166.2737169} are also open (even if not answered in this paper).
%%%% code-reuse
Finally, thanks to components, MSF improves code-reuse. Kernel components as well as any component (except the scheduler component which is specific to the simulation) can be reused from one simulation to another.

% no component-based framework for stencils orchestration (FFT, MapReduce/skeletons HLCM, nothing with DSLs)
To conclude this related work, as far as we know no component-based framework have already been proposed for stencils orchestration.