### 2017-02-22 ARTICLES/IJPP/articleR1/formalism.tex

Some approaches to numerically solve partial differential equations (PDEs) are based on direct iterative methods (\textit{e.g.} finite difference, finite volume or finite element methods).
They approximate the solution through a discretized process where the continuous time and space domains are discretized and numerical computations are iteratively (time discretization) applied onto a mesh (space discretization).
%In other words, the PDEs are transformed to a set of numerical computations applied at each time-step on elements of the discretized space domain.
While the computations can have various forms, we focus on three categories.
%TODO[JB] 4 plus tard...
\emph{Stencil} computations involve access to neighbor values only (the concept of neighborhood depending on the space discretization used).
\emph{Local} computations depend on the computed location only (this can be seen as a stencil of size one).
Finally, \emph{reductions} enable to transform values mapped on the mesh to a single scalar value.

This section gives a complete formal description of what we call a \textit{stencil program} and its computations. A Multi-Stencil Program meta-model is extracted from this formal description. The meta-model will be used to define MSL and GA in the next section.

%-------------------------------------
\subsection{Time, mesh and data}

$\Omega$ is the continuous space domain of a numerical simulation (typically $\mathbb{R}^n$). %For example, if $n=3$, any geographic point on earth could be represented in this space in a given geographic coordinate system.
A mesh $\mathcal{M}$ defines the discretization of the continuous space domain $\Omega$ of a set of PDEs and is defined as follows.

\begin{mydef}
A mesh is a connected undirected graph $\mathcal{M}=(V,E)$, where $V\subset \Omega$ is the set of vertices and $E\subseteq V^2$ the set of edges. The set of edges $E$ of a mesh $\mathcal{M}=(V,E)$ does not contain bridges. It is said that the mesh is applied onto $\Omega$.
\end{mydef}
\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{./images/maillages.pdf}}
  \caption{From left to right, Cartesian, curvilinear and unstructured meshes.}
  \label{fig:mesh1}
\end{center}\end{figure}
\begin{mydef}
The dimension of a mesh $\mathcal{M}=(V,E)$ applied onto $\Omega=\mathbb{R}^n$ is denoted $dim(\mathcal{M})=n$.
\end{mydef}
A mesh can be structured (as Cartesian or curvilinear meshes), unstructured, regular or irregular (without the same topology for each element) as illustrated in Figure~\ref{fig:mesh1}. One can notice that more than one type of mesh is possible inside a single simulation. For example, an hybrid mesh can be defined as an unstructured mesh composed itself of a Cartesian mesh inside each of its vertices. However, in this paper single mesh simulations are addressed.

\medskip

\begin{mydefs}[Definitions (mesh)]
\item An \textit{entity} $\phi$ of a mesh $\mathcal{M}=(V,E)$ is defined as a subset of its vertices and edges, $\phi\subset V\cup E$.
\item A \textit{group of mesh entities} $\mathcal{G} \in \mathcal{P}(V\cup E)$ represents a set of entities of the same topology and the mesh of a group is denoted $mesh(\mathcal{G})$, \textit{i.e.} $\mathcal{G} \in \mathcal{P}(V\cup E) \Leftrightarrow mesh(\mathcal{G})=(V,E)$
\item The \textit{set of entities groups} used in a simulation is denoted $\Phi$.
\end{mydefs}

For example, in a 2D Cartesian mesh, a group of entities where each entity is made of four vertices and four edges represents the cells.
Another group of entities are the vertices defined as all singletons formed of a single vertex of $V$. Both groups, then, would be part of $\Phi$. This example is illustrated in Figure~\ref{fig:meshbase}.

\medskip

\begin{mydef}
The finite sequence $T: (t_n)_{n\in\llbracket 0, T_{max} \rrbracket}$ represents the discretization of the continuous time domain $\mathcal{T}=\mathbb{R}$.
To each discrete time-step $n\in\llbracket 0, T_{max} \rrbracket$, a time value $t_n\in\mathcal{T}$ is associated. The time step is $t_n-t_{n-1}$.
\end{mydef}

The time discretization can be as simple as a constant time-step with a fixed number of steps.
The time-step and the number of steps can also change on the fly during the simulation depending on variables values (see definitions below).

\medskip

\begin{mydefs}[Definitions (quantity)]
\item $\mathbb{V}=\Delta\cup\mathcal{S}$ is the set of \emph{variables} or \emph{quantities}.
\item $\Delta$ are the \textit{mesh variables}. A mesh variable $\delta \in \Delta$ associates to each pair of entity and time-step a value $\delta: \mathcal{G}\times T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
\item The group of entities a variable is mapped on is denoted $entity(\delta)=\mathcal{G}$.
\item $\mathcal{S}$ are the \textit{scalar variables}. A scalar variable $s \in \mathcal{S}$ associates to each time-step a value $s: T\mapsto \mathcal{V}_\delta$ where $\mathcal{V}_{\delta}$ is a value type.
\item Among the scalar variables is one specific variable $conv\in\mathcal{S}$, the convergence criteria, whose value is $0$ except at the last step where it is 1. Thus, $\forall t\in \llbracket 0, T_{max-1} \rrbracket$, $conv(t)=0$, and $conv(T_{max})=1$.
\end{mydefs}

\medskip

This section has presented the formalism of meshes, their associated entities, groups of entities,  and also time discretization and quantities (mapped on meshes or scalars).

%-----------------------
\subsection{Computations}

\medskip
\begin{mydefs}
\item A computation domain $D$ is a subpart of a group of mesh entities, $D \subseteq \mathcal{G} \in \Phi$.
\item The set of computation domains of a numerical simulation is denoted $\mathcal{D}$.
\item $\mathcal{N}$ are the neighborhood functions $n: \mathcal{G}_i \mapsto \mathcal{G}_j^m$ which for a given entity $\phi \in \mathcal{G}_i$ returns a set of $m$ entities in $\mathcal{G}_j$. One can notice that $i = j$ is possible. Most of the time, such a neighborhood is called a \emph{stencil shape}.
\end{mydefs}

\begin{mydef}
A computation kernel $k$ of a numerical simulation is defined as $k=(S,R,(w,D),comp)$, where
\begin{itemize}
\item $S \in \mathcal{S}$ is the set of scalar to read,
\item $(w,D) \in \mathbb{V} \times \mathcal{D}$ is the single data written by the kernel:
\begin{itemize}
  \item $w$ is the single quantity (variable) modified by the computation kernel,
  \item $D$ is the computation domain on which $w$ is computed, $D \subseteq entity(w)$,
\end{itemize}
\item $R \in \Delta \times \mathcal{N}$ is the set of tuples $(r,n)$ representing the data read where
\begin{itemize}
  \item $r$ is a mesh variable read by the kernel to compute $w$,
  \item $n : entity(w) \rightarrow entity(r)^m$ is a neighborhood function that indicates which values of $r$ are read to compute $w$.
\end{itemize}
\item $comp$ is the numerical computation which returns a value from a set of $n$ input values, $comp: \mathcal{V}_i^n \rightarrow \mathcal{V}_j$, where $\mathcal{V}_i$ and $\mathcal{V}_j$ are value types. Thus, $comp$ represents the actual numerical expression which is computed by a kernel.
\end{itemize}
\end{mydef}

In a Multi-Stencil simulation, at each time-step, a set of computations is performed. During a computation kernel, it can be considered that a set of old states ($t-1$) of quantities are read ($\mathcal{S}$ and $R$), and that a new state ($t$) of a single quantity is written ($w$).

Such a definition of a computation kernel covers a large panel of different computations. For example, the four usual types of computations (stencil, local, boundary and reduction) performed into a simulation can be defined as follow :
\begin{itemize}
\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{stencil kernel} if $\exists (r,n) \in R$ such that $n \neq identity$ .
\item A \emph{boundary kernel} is a kernel $k(S,R,(w,D),comp)$ where $D$ is a specific computation domain at the border of entities, and which does not intersect with any other computation domain.
\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{local kernel} if $\forall (r,n) \in R$, $n = identity$.
\item A kernel computation $k(S,R,(w,D),comp)$ is a \emph{reduction kernel} if $w$ is a scalar. A reduction is typically used to compute the convergence criteria of the time loop of the simulation.
\end{itemize}

Since we only consider explicit numerical schemes in this paper, a kernel cannot write the same quantity it reads, except locally, \textit{i.e.} if $\exists (w,n)$ where $w \in R\Rightarrow n=identity$.

At this point, it could seem counter-intuitive to restrict the definition of a computation to a single quantity to write. This is actually counter-intuitive for a computer scientist who will naturally perform more than one computation into a loop to optimize cache memory use and to get a shorter code. The definition used above actually follows the usual "=" meaning of an equation or a numerical computation, where a single variable is written on the left of the symbol. We have made this choice for three reasons.
\begin{itemize}
\item First, we think this notation is more natural for a numerician who usually manipulate equations.
\item Second, we call a \emph{fusion} the fact of writing more than one quantity into a single loop. When writing a sequential program, fusion is natural for the developer, however, writing a parallel code without knowing it (as it is by using MSF), fusion are error prone and a bad fusion could lead to an incorrect numerical result.
\item Finally, when performing task parallelism, as it is in this paper, it is not always the good choice to perform fusions as you reduce the number of possible tasks.
\end{itemize}
For these reasons we have made the choice to restrict the computation definition without any fusion, and to propose an automatic loop fusion described in Section~\ref{sect:parallelism}. Moreover, one can note that the usual \emph{scatter} optimization of numerical codes is actually a specific case of fusion.

\begin{mydef}
The set of $n$ ordered computation kernels of a numerical simulation is denoted $\Gamma = [k_i]_{0 \leq i \leq n-1}$, such that $\forall k_i,k_j \in \Gamma$, if $i < j$, then $k_i$ is computed before $k_j$.
\end{mydef}

\begin{mydef}
A \textit{multi-stencil program} is defined by the octuplet 
\begin{equation}
\mathcal{MSP}(\mathcal{M},\Phi,\mathcal{D},\mathcal{N},\Delta, \mathcal{S},T,\Gamma)
\label{eq:msp}
\end{equation}
\end{mydef}

\paragraph{\textbf{Example}}
\begin{figure}[ht]
\begin{center}
\subfloat[Mesh and two groups of mesh entities.\label{fig:meshbase}]{
\resizebox{8cm}{!}{\includegraphics{./images/mesh.pdf}}
}\\
\hspace{10pt}
\subfloat[A is computed with a 4-neighborhood stencil applied on B. A is computed onto a computation domain which does not include all entities of the group.\label{fig:ex1}]{
\resizebox{5cm}{!}{\includegraphics{./images/stencil1.pdf}}
}
\vspace{20pt}
\hspace{10 pt}
\subfloat[A is computed with a 2-neighborhood stencil applied on C. A is computed onto a computation domain includes all entities of the group.\label{fig:ex2}]{
\resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
}
\end{center}
\caption{(a) a Cartesian mesh and two kind of groups of mesh entities, (b) an example of stencil kernel on cells, (c) an example of stencil kernel on two different entities of the mesh.}
\label{fig:gspmsp}
\end{figure}
For example, in Figure~\ref{fig:ex1}, assuming that the computation domain (full lines) is denoted $dc1$ and the stencil shape is $n1$, the stencil kernel can be defined as:
\begin{equation*}
R: \{(B,n1)\}, \quad w: A, \quad D: dc1,
\end{equation*}
\begin{equation*}
comp: A(x,y)=B(x+1,y)+B(x-1,y)+B(x,y+1)+B(x,y-1).
\end{equation*}
On the other hand, in the example of Figure~\ref{fig:ex2}, assuming the computation domain is $dc2$ and the stencil shape is $n2$, the stencil kernel is defined as:
\begin{equation*}
R: \{(C,n2),(A,identity)\}, \quad w: A, \quad D: dc2,
\end{equation*}
\begin{equation*}
comp: A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1).
\end{equation*}

\medskip
A stencil program has been formally defined in this section. This formalism is used in the next Section as a meta-model to define both a the domain specific language MSL, and the generic assembly of a multi-stencil program GA.

One can note that the meta-definition(~\ref{eq:msp}) is independent from the topology of the mesh. This property will be kept in the rest of this paper so that MSF has the capacity to be used for any kind of mesh, thus being mesh-agnostic.