### 2017-03-03 ARTICLES/IJPP/articleR1/article.tex
abstract + titre

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
%----------------------------------------
\input{header.tex}
%----------------------------------------

%----------------------------------------
\begin{document}
%----------------------------------------

%----------------------------------------
\title{The Multi-Stencil Framework: A Component-Based Approach to Orchestrate Stencils}
%----------------------------------------
\author{H\'el\`ene Coullon \and Julien Bigot \and Christian Perez}
%----------------------------------------
\institute{H\'el\`ene Coullon \at
              DAPI IMT Atlantique, LS2N, Inria. Nantes, France\\
              \email{helene.coullon@inria.fr}
	\and
           Julien Bigot \at
           Maison de la Simulation, CEA, CNRS, Univ. Paris-Sud, UVSQ, Universit\'e Paris-Saclay, 91191 Gif-sur-Yvette, France\\
           \email{julien.bigot@cea.fr}           
           \and
	Christian Perez \at
              Univ. Lyon, Inria, CNRS, ENS de Lyon. Lyon, France\\
              \email{christian.perez@inria.fr}
}
%----------------------------------------
\maketitle
%----------------------------------------
\begin{abstract}
As the computation power of modern high performance architectures increases, their heterogeneity and complexity also become more important. One of the big challenges of exascale is to reach programming models that give access to high performance computing (HPC) to many scientists and not only to a few HPC specialists. One relevant solution to ease parallel programming for scientists is Domain Specific Language (DSL). However, one problem to avoid with DSLs is to mutualize existing codes and libraries instead of implemented from scratch. For example, this phenomenon occurs for stencil-based numerical simulations, for which a large number of languages has been proposed without code reuse between them. 
The Multi-Stencil Framework (MSF) presented in this paper combines a meta-model to component-based programming models to enhance code reuse and separation of concerns in the specific case of stencils. MSF can easily choose one parallelization technique or another as well as one optimization or another, as well as one back-end implementation or another. It is shown that MSF can reach same performances than a non-component-based MPI implementation over 16.384 cores. Finally, the performance model of the framework for hybrid parallelization is validated by evaluations.

\keywords{Component programming models \and Domain Specific Language (DSL) \and Stencil \and Numerical simulation \and Data parallelism \and Task parallelism \and Scheduling \and MPI \and OpenMP}
\end{abstract}
%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
\input{introduction.tex}
%------------------------------------------------------------------------------
\section{The Component-Based Multi-Stencil Framework}
\label{sect:msf}
\input{msf.tex}
%------------------------------------------------------------------------------
\section{Meta-modelization of a Multi-Stencil Program}
\label{sect:formalism}
\input{formalism.tex}
%------------------------------------------------------------------------------
\section{Generic Asssembly and The Multi-Stencil Language}
\label{sect:msl}
\input{msl.tex}
%------------------------------------------------------------------------------
\section{The Multi-Stencil Compiler}
\label{sect:parallelism}
\input{parallelism.tex}
%------------------------------------------------------------------------------
%\section{Static Scheduling and performance model}
%\label{sect:msp}
%\input{msp.tex}
%------------------------------------------------------------------------------
%\section{Implementation}
%\label{sect:comp}
%\input{compiler.tex}
%------------------------------------------------------------------------------
\section{Evaluation}
\label{sect:eval}
\input{evaluation.tex}
%------------------------------------------------------------------------------
\section{Related work}
\label{sect:rel}
\input{related.tex}
%------------------------------------------------------------------------------
\section{Conclusion}
\label{sect:concl}
\input{conclusion.tex}
%------------------------------------------------------------------------------
\begin{acknowledgements}
This work has partially been supported by the PIA ELCI project of the French FSN.
This  work  was  granted  access  to  the  HPC  resources  of TGCC under  the 
allocations t2015067470, x2016067617 and AP010610191 made by GENCI.
\end{acknowledgements}
%----------------------------------------
\bibliographystyle{plain}
\bibliography{article}
%----------------------------------------
\end{document}
%----------------------------------------