### 2017-03-06 ARTICLES/IJPP/articleR1/msl.tex
corrections formalism.tex + msl.tex

From the octuplet of Equation~(\ref{eq:msp}) both the Generic Assembly (GA) of a multi-stencil program and the Multi-Stencil Language (MSL) can be built. GA and MSL are both are described in this section.

%----------------------------------------
\subsection{Generic Assembly}
%----------------------------------------
As illustrated in Figure~\ref{fig:ga}, the GA has five components: {\bf \emph{Driver}, \emph{Time}, \emph{DDS}, \emph{Data}, } and {\bf \emph{Computations}}. These components are generic components or abstract components. It means that interfaces of these components are well defined but that they are not implemented yet in GA. They can be compared to abstract classes and templates in \texttt{C++} for which an implementation must be given as well as specific parameters.

\begin{figure}[t]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component,fill=green!50!white] (D) at (0,0) {$Driver$};
   \node[provide] (Dp) at (-1,0) {};
   \node (Ds) at (-1.5,0) {start};
   \node[use,right=1.5cm of D] (Du1) {};
   \node[use,below=1.75cm of D] (Du2) {};
   \node[use,below=0.8cm of Du1] (Du3) {$m$};

   \node[provide,below=0.15 of Du2] (Tp) {};
   \node[component,below=1.6cm of Tp,fill=green!50!white] (T) {$Time$};
   \node[use,right=1cm of T] (Tu) {};
   \node[left=0.8cm of T] (tt) {$T$};

   \node[provide,right=0.15 of Tu] (Cp) {};
   \node[component, draw=blue!60!white,fill=green!50!white,text=blue!80!black,line width=0.5mm,right=2cm of Cp] (C) {$Computations$};
   \node[use,above=0.8cm of C] (Cu) {};
   \node[right=0.2cm of Cu] (star) {$*$};
   \node[right=1.5cm of C] (gamma) {$\Gamma$};

   \node[provide,below right=0.2 of Du3] (Datap1) {};
   \node[provide,above=0.15 of Cu] (Datap2) {};
   \node[component,double, draw=red,fill=red!50!white,above=0.8cm of Datap2] (Data) {$Data$};
   \node[use,above=0.8cm of Data] (Datau) {};
   \node[left=1cm of Data] (delta) {$\Delta$};
   \node[mpi,right=1cm of Data] (mpidata) {};
   \node[above=0.2cm of mpidata] (star2) {$*$};

   \node[provide,right=0.15 of Du1] (DDSp1) {};
   \node[provide,above=0.15 of Datau] (DDSp2) {};
   \node[component, draw=red, fill=red!50!white,above=0.8cm of DDSp2] (DDS) {$DDS$};
   \node[right=1.5cm of DDS] (m) {$\mathcal{M},\Phi,\mathcal{D},\mathcal{N}$};

  \path[-]
    (Dp) edge node {} (D)
    (D) edge node {} (Du1)
        edge node {} (Du2)
        edge node {} (Du3)
    (DDSp1) edge node {} (DDS)
    (Tp) edge node {} (T)
    (T)  edge node {} (Tu)
    (Cp) edge node {} (C)
    (C) edge node {} (Cu)
    (Datap2) edge node {} (Data)
    (Datap1) edge node {} (Data)
    (Data) edge node {} (Datau)
          edge node {} (mpidata)
    (DDSp2) edge node {} (DDS);
\end{tikzpicture}
\vspace*{.5em}
\caption{Generic Assembly according to the Multi-Stencil program formalism.
Components circled by a double line identify those that will be instantiated multiple times by MSC. Component colors represent actors of Figure~\ref{fig:msf} responsible for the component implementation: green for those automatically generated by the compiler, red for those implemented by HPC specialists and blue for those implemented by the developer.}
\label{fig:ga}
\end{center}
\end{figure}

\paragraph{\textbf{Driver}}
This component can be compared to the main function of a usual program. It is responsible for both the initialization and the execution of other components (like variable initialization and function calls). This component is generated by MSC (represented in green).

\paragraph{\textbf{Time}}
This component is responsible for the time $T$ defined in Equation~(\ref{eq:msp}). It is composed of a time loop and potentially of a convergence reduction. This component is generated by MSC (represented in green).

\paragraph{\textbf{DDS}}
This component is responsible for the mesh and its entities $\mathcal{M}$ and $\Phi$, the set of computation domains $\mathcal{D}$, and the set of neighborhood functions $\mathcal{N}$.
When the generic assembly is instantiated and specialized by MSC, an implementation of DDS is selected to handle a specific type of mesh.
%\CP{pas clair}
The interfaces exposed by this component are well defined and any component providing these interfaces can be indifferently used.
A third party specialist can therefore propose new implementation of DDS.
In this paper, both data and task parallelism are used. In the case of data parallelism DDS handles mesh partitioning and provides a synchronization interface as detailed in Section~\ref{sect:parallelism}. The implementation of this component is the responsibility of HPC specialists (represented in red).

\paragraph{\textbf{Data}}
This component is responsible for the set of mesh variables $\Delta$.
Each instance of the component uses the DDS component to handle one single mesh variable.
It is closely related to $DDS$ and its implementation is typically provided by the same HPC specialist as $DDS$ (represented in red).

\paragraph{\textbf{Computations}}
This component is responsible for $\Gamma$, \ie the computations of the simulation. 
It is automatically replaced by a sub-assembly of components produced by MSC for which the parallel part is automatically generated. On the other hand, components responsible for the numerical kernels are filled by the developer. This is why this component is represented in green and blue in Fig.~\ref{fig:ga}. The sub-assembly generation is described in Section~\ref{sect:parallelism}.

%----------------------------------------
\subsection{The Multi-Stencil Language}
%----------------------------------------

The second element of MSF which is built upon the meta-model represented by Equation~(\ref{eq:msp}) is the Multi-Stencil Language and its grammar. This grammar is light and descriptive only. However it is sufficient (in addition to GA) for MSC to automatically extract a parallel pattern of the simulation, which is finally dumped as a specialized instantiation of GA.

\begin{filecontents*}{grammar.txt}
program ::= "mesh:" meshid 
            "mesh entities:" listgroup
            "computation domains:" listcompdom
            "independent:" listinde
            "stencil shapes:" liststencil
            "mesh quantities:" listquantities
            "scalars:" listscalar
            listloop

listgroup ::= groupid "," listgroup | groupid
listcompdom ::= compdom listcompdom | compdom
compdom ::= compdomid "in" groupid
listinde ::= inde listinde | inde
inde ::= compdomid "and" compdomid
liststencil ::= stencil liststencil | stencil
stencil ::= stencilid "from" groupid "to" groupid
listquantities ::= quantity listquantities |  quantity
quantity ::= groupid listquantityid
listquantityid ::= quantityid "," listquantityid |  quantityid
listscalar ::= scalarid "," listscalar | scalarid
listloop ::= loop listloop | loop
loop ::=  "time:" iteration
          "computations:" listcomp
iteration ::= num | scalarid
listcomp ::= comp listcomp |  comp
comp ::= written "=" compid "(" listread ")"
written ::= quantityid "[" compdomid "]" | scalar
listread ::= dataread listread | dataread
dataread ::= quantityid "[" stencilid "]" |  quantityid | scalar
\end{filecontents*}

\begin{figure}[hbt]
  \hspace{5mm}
  \begin{minipage}[!h]{0.98\textwidth}
    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left]{grammar.txt}}   
    \caption{Grammar of the Multi-Stencil Language. \label{fig:grammar}}
  \end{minipage}
\end{figure}

The grammar of the Multi-Stencil Language is given in Figure~\ref{fig:grammar} and an example is provided in Figure~\ref{fig:exmsl}. A Multi-Stencil program is composed of eight parts that match those of Equation~(\ref{eq:msp}).

\begin{filecontents*}{exmsl.txt}
mesh : cart
mesh entities : cell, edgex
computation domains :
  d1 in cell
  d2 in edgex
independent :
  d1 and d2
stencil shapes : 
  ncc from cell to cell
  nce from cell to edgex
  nec from edgex to cell
mesh quantities :
  cell A,B,D,E,F,G,I,J
  edgex C,H
scalars : mu, tau
time : 500
computations :
  B[d1] = k0(tau,A)
  C[d2] = k1(B[nec])
  D[d1] = k2(C)
  E[d1] = k3(C)
  F[d1] = k4(D,C[nce])
  G[d1] = k5(mu,tau,E)
  H[d2] = k6(F)
  I[d1] = k7(G,H)
  J[d1] = k8(mu,I[ncc])
\end{filecontents*}

\begin{figure}[!h]
  \hspace{5mm}
  \begin{minipage}[!h]{0.98\textwidth}
    {\lstinputlisting[basicstyle=\small,mathescape,frame=single,language=Python,numbers=left]{exmsl.txt}}   
    \caption{Example of program using the Multi-Stencil Language. \label{fig:exmsl}}
  \end{minipage}
\end{figure}

\begin{enumerate}

\item The \texttt{mesh} keyword (Fig.~\ref{fig:grammar},~l.1) introduces an identifier for $\mathcal{M}$, the single mesh of the simulation. For example \texttt{cart} in Fig.~\ref{fig:exmsl},~l.1. The language, based on the meta-model is independent of the mesh topology, thus this identifier is actually not used by the compiler.% but is syntactic sugar for the user.
%\JB{J'aime pas trop l'utilisation de ``syntactic sugar'' ici... Syntactic sugar est generalement utilise pour un racourci de syntaxe (i.e. int x = myInteger au lieu de int x = myInteger.getIntVal()), ce n'est pas le cas ici}

\item The \texttt{mesh entities} keyword (Fig.~\ref{fig:grammar},~l.2) introduces identifiers for the groups of mesh entities $\mathcal{G}\in\Phi$. For example \texttt{cell} or \texttt{edgex} in Fig.~\ref{fig:exmsl},~l.2. %Again, this is syntactic sugar only as the compiler does not rely on the mesh topology.

\item The \texttt{computation domains} keyword (Fig.~\ref{fig:grammar},~l.3) introduces identifiers for the computation domains $D\in\mathcal{D}$. For example \texttt{d1} and \texttt{d2} in Fig.~\ref{fig:exmsl},~l.4-5. For reference, each domain is associated to a group of entities (Fig.~\ref{fig:grammar},~l.12) such as \texttt{cell} for \texttt{d1} in Fig.~\ref{fig:exmsl},~l.4.

\item The \texttt{independent} keyword (Fig.~\ref{fig:grammar},~l.4) offers a way to declare that computation domains do not intersect, such as \texttt{d1} and \texttt{d2} in  Fig.~\ref{fig:exmsl},~l.7. This is used by the compiler to compute dependencies between computations.

\item The \texttt{stencil shapes} keyword (Fig.~\ref{fig:grammar},~l.5) introduces identifiers for each stencil shape $n\in\mathcal{N}$. For each $n$, the source and destination group of entities (Fig.~\ref{fig:grammar},~l.16) are specified. For example \texttt{nec} in Fig.~\ref{fig:exmsl},~l.11 is a neighborhood from \texttt{edgex} to \texttt{cell}.

\item The \texttt{mesh quantities} keyword (Fig.~\ref{fig:grammar},~l.6) introduces identifiers for $\delta\in\Delta$, the mesh variables with the group of entities they are mapped on (Fig.~\ref{fig:grammar},~l.16). For example the quantities \texttt{C} and \texttt{H} are mapped onto the groups of mesh entities \texttt{edgex}.

\item The \texttt{scalars} keyword (Fig.~\ref{fig:grammar},~l.7) introduces identifiers for $s\in\mathcal{S}$, the scalars. For example \texttt{mu} and \texttt{tau} in Fig.~\ref{fig:exmsl},~l.15.

\item Finally, the last part (Fig.~\ref{fig:grammar},~l.8) introduces the different computation loops of the simulation. Each loop is made of two parts:
\begin{itemize}
\item the \texttt{time} keyword (Fig.~\ref{fig:grammar},~l.22) introduces either a constant number of iterations or $conv$, the convergence criteria that is a scalar (Fig.~\ref{fig:grammar},~l.24). For example, 500 iterations are specified in Fig.~\ref{fig:exmsl},~l.16,
\item the \texttt{computations} keyword (Fig.~\ref{fig:grammar},~l.23) introduces identifiers for each computation $k=(S,R,(w,D),comp) \in \Gamma$. Each computation (Fig.~\ref{fig:grammar},~l.26) specifies:
\begin{itemize}
 \item the quantity $w$ written and its domain $D$, for example in Fig.~\ref{fig:exmsl},~l.22, kernel \texttt{k4} computes the variable \texttt{F} on domain \texttt{d1},
 \item the read scalars $S$ and mesh variables with their associated stencil shape ($R$). For example in Fig.~\ref{fig:exmsl},~l.16, \texttt{k4} reads \texttt{C} with the shape \texttt{nce} and \texttt{D} with the default \textit{identity} shape; it does not read scalars.
 \end{itemize}
\end{itemize}
\end{enumerate}

One can notice that in the example of Figure~\ref{fig:exmsl}, there are no kernel associated to the scalars \texttt{mu} and \texttt{tau} (reduction).
In this case, those scalars are in fact constants.
One can also notice that the computation to execute for each kernel is not specified. Only an identifier is given to each kernel, for example $k_4$. The numerical code is indeed not handled by MSL that generates a parallel orchestration of computations only. The numerical computation is specified after MSC compilation by the developer (Fig.~\ref{fig:msf}).